### 2024-04-05 code/analysis-code/fullanalysis-v1.qmd
Project Part 4
Completed full analysis and updated all files.

---
title: "Full Analysis Script"
author: "Cassia Roth"
date: "`r Sys.Date()`"
output: html_document
---

# Setup

```{r, message=FALSE,  warning=FALSE}
#load needed packages. make sure they are installed.
library(tidyverse) #for data processing/cleaning; includes ggplot2, tidyr, readr, dplyr, stringr, purr, forcats
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(renv) #for package management
library(knitr) #for nice tables
library(kableExtra) #for nice tables
library(gt) #for nice tables
library(gtsummary) #for summary tables
library(ggplot2) #for plotting
library(tidymodels) #for modeling

rm(list=ls()) #clear the environment
```

Load the data

```{r, message=FALSE,  warning=FALSE}
#Path to summary data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","ML_linear.rds")
#load data
ML_linear <- readRDS(data_location)
```

The first simple linear regression model is as follows:

$$\text{BW} = \alpha + \beta_1 \text{MA} + \epsilon$$

whereas BW is birthweight in grams, MA is maternal ancestry (Euro-descent, Afro-descent), and $\epsilon$ is the error term. Results are displayed in Table 2.

```{r}
# Model using recipes and workflows
# Define recipe for model
ML_recipe1 <- recipe(Weightgrams ~ ModifiedColor, data = ML_linear) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable

# Specify model
ML_spec1 <- linear_reg() %>% 
  set_engine("lm")

# Create work flow
ML_wf1 <- workflow() %>%
  add_recipe(ML_recipe1) %>%
  add_model(ML_spec1)

# Fit the work flow
ML_fit1 <- fit(ML_wf1, data = ML_linear)

# Summarize model fit with tidiers
ML_summary1 <- summary(ML_fit1) # Summary of fit
tidy(ML_fit1, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit1) # Model-level statistics

# Produce predictions for R-squared
ML_fit1_pred <- predict(ML_fit1, new_data = ML_linear %>% select(-Weightgrams))
ML_fit1_pred

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit1_pred <- bind_cols(ML_fit1_pred, ML_linear %>% select(Weightgrams))
ML_fit1_pred

# Plot observed versus predicted
plot1 <- ggplot(ML_fit1_pred, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot1

# Create metric set including RMSE and R-squared
metrics_fit1 <- metric_set(rmse, rsq)
metrics_fit1(ML_fit1_pred, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit1 <- as.data.frame(metrics_fit1(ML_fit1_pred, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit1

# Create a gtsummary table directly from the model
table2 <-
  tbl_regression(ML_fit1, exponentiate = FALSE, intercept = TRUE, show_single_row = everything()) %>% 
  bold_p() %>% 
  add_glance_table(include = c(r.squared, adj.r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE)
table2

#Print table
table2

# Save regression output into table
saveRDS(table2, file = here::here("results", "tables", "table2_final.rds"))
```

The second multiple linear regression model is as follows:

$$\text{BW} = \alpha + \beta_1 \text{MA} + \beta_2 \text{Gest} + \beta_3 \text{Nat} + \beta_4 \text{Age} + \epsilon$$

whereas BW is birthweight in grams, MA is maternal ancestry (Euro-descent, Afro-descent), Gest is gestational status (nullipara, primipara, multipara, multigravida), Age is maternal age in years, and $\epsilon$ is the error term. Results are displayed in Table 3.

```{r}
# Using recipes and workflows
# Define recipe for model
ML_recipe2 <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Define model specification
ML_spec2 <- linear_reg() %>% 
  set_engine("lm") %>%
  set_mode("regression")

# Create workflow
ML_wf2 <- workflow() %>%
  add_recipe(ML_recipe2) %>%
  add_model(ML_spec2)

# Fit the workflow
ML_fit2 <- fit(ML_wf2, data = ML_linear)

# Summarize model fit with tidiers
ML_lm2_summary <- summary(ML_fit2) # Summary of fit
tidy(ML_fit2, conf.int = TRUE) %>% mutate(p.value = round(p.value, 5)) # Coefficients and confidence intervals with p-values rounded to 5 decimal places
glance(ML_fit2) # Model-level statistics

# Create a gtsummary table directly from the model
table3 <- pull_workflow_fit(ML_fit2) %>%
  tbl_regression(exponentiate = FALSE, intercept = TRUE) %>%
  add_glance_table(include = c(r.squared, adj.r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE) %>%
  bold_p()
table3

# Save regression output into table
saveRDS(table3, file = here::here("results", "tables", "table3_final.rds"))
```

The second two models are also simple and multiple linear regressions with outcome variable as birthweight in grams. However, in these two models, I maintained maternal skin color as three categories, White (the reference group), Black, and mixed race to see if there is differential outcomes for Black and mixed-race women. The first simple linear regression model is as follows:

$$\text{BW} = \alpha + \beta_1 \text{MC} + \epsilon$$
whereas BW is birthweight in grams, MC is maternal skin color (White, Black, and mixed race), and $\epsilon$ is the error term. Results are displayed in Table 4.

```{r}
# Using recipes and workflows
# Simple linear model
# Specify recipe
ML_recipe3 <- recipe(Weightgrams ~ Color, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White")

# Specify model
ML_spec3 <- linear_reg() %>%
  set_engine("lm")

# Create workflow
ML_wf3 <- workflow() %>%
  add_recipe(ML_recipe3) %>%
  add_model(ML_spec3)

# Fit the workflow
ML_fit3 <- fit(ML_wf3, data = ML_linear)

# Summarize model fit with tidiers and create a gtsummary table directly from the model
ML_summary3 <- summary(ML_fit3)
tidy(ML_fit3, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit3) # Model-level statistics

# Create a gtsummary table from model
table4 <- tbl_regression(ML_fit3, exponentiate = FALSE, intercept = TRUE) %>%
  bold_p() %>%
  add_glance_table(include = c(r.squared, adj.r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE)
table4

# Save regression output into table
saveRDS(table4, file = here("results", "tables", "table4_final.rds"))
```

The second multiple linear regression model is as follows:

$$\text{BW} = \alpha + \beta_1 \text{MC} + \beta_2 \text{Gest} + \beta_3 \text{Nat} + \beta_4 \text{Age} + \epsilon$$

whereas BW is birthweight in grams, MC is maternal skin color (White, Black and mixed race), Gest is gestational status (nullipara, primipara, multipara, multigravida), Nat is maternal nationality (Brazilian, Latin American, European, Middle Eastern), Age is maternal age in years, and $\epsilon$ is the error term.

```{r}
# Using reipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4 <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multipara") %>% # Set reference level for ModifiedStatus
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Specify model
ML_spec4 <- linear_reg() %>%
  set_engine("lm")

# Create workflow
ML_wf4 <- workflow() %>%
  add_recipe(ML_recipe4) %>%
  add_model(ML_spec4)

# Fit the workflow
ML_fit4 <- fit(ML_wf4, data = ML_linear)

# Summarize model fit with tidiers and create a gtsummary table directly from the model
ML_summary4 <- summary(ML_fit4)
tidy(ML_fit4, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit4) # Model-level statistics

# Create a gtsummary table from model
table5 <- tbl_regression(ML_fit4, exponentiate = FALSE, intercept = TRUE) %>%
  bold_p() %>%
  add_glance_table(include = c(r.squared, adj.r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE)
table5

# Save regression output into table
saveRDS(table5, file = here("results", "tables", "table5_final.rds"))
```