### 2024-05-01 project-review-by CLi.qmd

---
title: Project Part 5 Review  
author: Chaohua Li
date:  04-25-2024
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project:Low Infant Birth Weight in Brazil: A Historical Data Analysis Approach

Name of project author(s):  Cassia Roth

Name of project reviewer:   Chaohua Li

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The introduction is well-structured with citations appropriately added. The selection of the research question is justified.

### Summary assessment (PICK ONE, DELETE THE OTHERS)

* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The question was clearly stated.

### Summary assessment

* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

Data source was described in detail. A codebook was also provided in the 'rawdata' folder.

### Summary assessment

* source and overall structure of data well explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

Data was properly cleaned and processed. Comments/Notes were provided for each procedure.

### Summary assessment

* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

The analysis methods were appropriate given the inferential purpose of this study. Model performances were compared using cross-validation. Final model was evaluated using the test data.

### Summary assessment

* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

The results were presented in a well-structured manner. It was easy to follow the steps without getting lost. The format for Tables 2&3&5 may need a little tweak. The first column and the last column of the tables appear to be truncated.

### Summary assessment

* results are very well presented

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

Findings were properly discussed. Strengths and limitations were acknowledged.

### Summary assessment

* strong, complete and clear discussion

## Further comments

The content of the paper was super clear and easy to follow. The only thing that needs improvement is the format of the tables.

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

All files are well label. All residual files from the template have been removed.

### Summary assessment

* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

There are Rmd files in each folder, clearly indicating the content and purpose for the files in it.

### Summary assessment

* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

I was able reproduce the results. Instructions were clear.

### Summary assessment

* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

The research questions have been fully addressed. This study is relatively complete and thorough.

### Summary assessment

* strong level of thorougness

## Further comments

Congratulations on completing the project so well! Enjoy your summertime.