### 2024-06-29 code/analysis-code/fullanalysis-v1.qmd
Merged GT tables; updated supplement
Revisions

---
title: "Full Analysis Script"
author: "Cassia Roth"
date: "`r Sys.Date()`"
output: html_document
---

# Setup

```{r, message=FALSE,  warning=FALSE}
#load needed packages. make sure they are installed.
#install.packages("glue")
library(tidyverse) #for data processing/cleaning; includes ggplot2, tidyr, readr, dplyr, stringr, purr, forcats
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(renv) #for package management
library(knitr) #for nice tables
library(kableExtra) #for nice tables
library(gt) #for nice tables
library(gtsummary) #for summary tables
library(ggplot2) #for plotting
library(tidymodels) #for modeling
library(glue)
rm(list=ls()) #clear the environment
```

Load the data

```{r, message=FALSE,  warning=FALSE}
#Path to summary data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","ML_linear.rds")
#load data
ML_linear <- readRDS(data_location)
```

# Preparation for Model Assessment

First, I will create train and test data (75/25) to be able to evaluate the performance of our models. I will also set the seed to ensure reproducibility. For each model, I perform the following steps: 1) run the model on the train data; 2) perform cross-validation on the train data; 3) evaluate the model on the test data. The model outputs included in the final manuscript are on all the data.

```{r, message=FALSE,  warning=FALSE}
# Fix the random numbers by setting the seed 
# This enables the analysis to be reproducible when random numbers are used 
rngseed = 1234
set.seed(rngseed)

# Create training data
# Put 3/4 of the data into the training set 
data_split <- initial_split(ML_linear, prop = 3/4)

# Create data frames for the two sets:
train_data <- training(data_split)
test_data  <- testing(data_split)
```

# Model 1

The first simple linear regression model is as follows:

$$\text{BW} = \alpha + \beta_1 \text{MA} + \epsilon$$

whereas BW is birth weight in grams, MA is maternal ancestry (Euro-descent, Afro-descent), and $\epsilon$ is the error term.

Full results are displayed in Table 2.

```{r}
# Model 1: Final model for paper using all data
# Model using recipes and workflows
# Define recipe for model
ML_recipe1 <- recipe(Weightgrams ~ ModifiedColor, data = ML_linear) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") #%>% # Set reference level for ModifiedColor
  #step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable. Cannot do this if I am merging tables

# Specify model
ML_spec1 <- linear_reg() %>% 
  set_engine("lm")

# Create work flow
ML_wf1 <- workflow() %>%
  add_recipe(ML_recipe1) %>%
  add_model(ML_spec1)

# Fit the work flow
ML_fit1 <- fit(ML_wf1, data = ML_linear)

# Summarize model fit with tidiers
ML_summary1 <- summary(ML_fit1) # Summary of fit
tidy(ML_fit1, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit1) # Model-level statistics

# Produce predictions for R-squared
ML_fit1_pred <- predict(ML_fit1, new_data = ML_linear %>% select(-Weightgrams))
ML_fit1_pred

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit1_pred <- bind_cols(ML_fit1_pred, ML_linear %>% select(Weightgrams))
ML_fit1_pred

# Plot observed versus predicted
plot1 <- ggplot(ML_fit1_pred, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot1

# Create metric set including RMSE and R-squared
metrics_fit1 <- metric_set(rmse, rsq)
metrics_fit1(ML_fit1_pred, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit1 <- as.data.frame(metrics_fit1(ML_fit1_pred, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit1

# Create a gtsummary table directly from the model
table_model1 <-
  tbl_regression(ML_fit1, exponentiate = FALSE, intercept = TRUE) %>% 
  add_significance_stars(hide_ci = TRUE, hide_p = TRUE) %>%
  add_glance_table(include = c(r.squared, nobs))
table_model1

#Print table
table_model1

# Save regression output into table
saveRDS(table_model1, file = here::here("results", "tables", "table_model1_final.rds"))
```
## Model 1 Performance

First we will evaluate the model on our training data.

```{r}
# Training model 1
# Model using recipes and workflows
# Define recipe for model
ML_recipe1_train <- recipe(Weightgrams ~ ModifiedColor, data = train_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable

# Specify model
ML_spec1_train <- linear_reg() %>% 
  set_engine("lm")

# Create work flow
ML_wf1_train <- workflow() %>%
  add_recipe(ML_recipe1_train) %>%
  add_model(ML_spec1_train)

# Fit the work flow
ML_fit1_train <- fit(ML_wf1_train, data = train_data)

# Summarize model fit with tidiers
ML_summary1_train <- summary(ML_fit1_train) # Summary of fit
tidy(ML_fit1_train, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit1_train) # Model-level statistics

# Produce predictions for R-squared
ML_fit1_pred_train <- predict(ML_fit1_train, new_data = train_data %>% select(-Weightgrams))
ML_fit1_pred_train

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit1_pred_train <- bind_cols(ML_fit1_pred_train, train_data %>% select(Weightgrams))
ML_fit1_pred_train

# Plot observed versus predicted
plot1_train <- ggplot(ML_fit1_pred_train, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot1_train

# Create metric set including RMSE and R-squared
metrics_fit1_train <- metric_set(rmse, rsq)
metrics_fit1_train(ML_fit1_pred_train, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit1_train <- as.data.frame(metrics_fit1_train(ML_fit1_pred_train, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit1_train

# Create a gtsummary table directly from the model
table_model1_train <-
  tbl_regression(ML_fit1_train, exponentiate = FALSE, intercept = TRUE, show_single_row = everything()) %>% 
  bold_p() %>% 
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE)
table_model1_train

# Save regression output into table
saveRDS(table_model1_train, file = here::here("results", "tables", "table_model1_train.rds"))
```
Our results here are not that different from the model run on all the data. Our adjusted R-squared is 0.01, the same as the full model. The RMSE is 488, which is slightly lower than the full model (494), which shows better performance.

Now let's perform cross-validation on the training data.

```{r, message=FALSE,  warning=FALSE}
# CV model 1
# Set the seed for reproducibility
set.seed(rngseed)

# Create 10-fold cross-validation
folds <- vfold_cv(train_data, v = 10)
folds

# Define recipe for model
ML_recipe1_cv <- recipe(Weightgrams ~ ModifiedColor, data = train_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable

# Set model specification
ML_spec1_cv <- linear_reg() %>% set_engine("lm")

# Create and initialize workflow that bundles together model specification and formula.
ML_wf1_cv <- workflow() %>% 
  add_model(ML_spec1_cv) %>% 
  add_recipe(ML_recipe1_cv)
  fit_resamples(ML_wf1_cv, resamples = folds)

# Resample the model
ML_wf1_cv_rs <- fit_resamples(ML_wf1_cv, resamples = folds)
ML_wf1_cv_rs

# Get model metrics
ML_cv1_metrics <- collect_metrics(ML_wf1_cv_rs)

# Filter the data frame to include only rows with "rmse" and "rsq" in the ".metric" column
ML_cv1_metrics <- ML_cv1_metrics[ML_cv1_metrics$.metric %in% c("rmse", "rsq"), !(names(ML_cv1_metrics) %in% ".config")]
ML_cv1_metrics

# Save regression output into table
saveRDS(ML_cv1_metrics, file = here::here("results", "tables", "ML_cv1_metrics.rds"))
```
Performing cross-validation results in an RMSE that is the same as our training model (488), lower than our full model. The r-squared (0.015) is also higher than our training and full models (0.01), demonstrating better model performance.

Finally, let's run the model on our test data.

```{r, message=FALSE,  warning=FALSE}
# Testing model 1
# Model using recipes and workflows
# Define recipe for model
ML_recipe1_test <- recipe(Weightgrams ~ ModifiedColor, data = test_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable

# Specify model
ML_spec1_test <- linear_reg() %>% 
  set_engine("lm")

# Create work flow
ML_wf1_test <- workflow() %>%
  add_recipe(ML_recipe1_test) %>%
  add_model(ML_spec1_test)

# Fit the work flow
ML_fit1_test <- fit(ML_wf1_test, data = test_data)

# Summarize model fit with tidiers
ML_summary1_test <- summary(ML_fit1_test) # Summary of fit
tidy(ML_fit1_test, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit1_test) # Model-level statistics

# Produce predictions for R-squared
ML_fit1_pred_test <- predict(ML_fit1_test, new_data = test_data %>% select(-Weightgrams))
ML_fit1_pred_test

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit1_pred_test <- bind_cols(ML_fit1_pred_test, test_data %>% select(Weightgrams))
ML_fit1_pred_test

# Plot observed versus predicted
plot1_test <- ggplot(ML_fit1_pred_test, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot1_test

# Create metric set including RMSE and R-squared
metrics_fit1_test <- metric_set(rmse, rsq)
metrics_fit1_test(ML_fit1_pred_test, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit1_test <- as.data.frame(metrics_fit1_test(ML_fit1_pred_test, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit1_test

# Create a gtsummary table directly from the model
table_model1_test <-
  tbl_regression(ML_fit1_test, exponentiate = FALSE, intercept = TRUE, show_single_row = everything()) %>% 
  bold_p() %>% 
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE)
table_model1_test

# Save regression output into table
saveRDS(table_model1_test, file = here::here("results", "tables", "table_model1_test.rds"))
```

When running the model on our testing dataset, our results remain significant at the 0.05 level. The RMSE is 508 and the r-squared is 0.008. This shows a slightly worse fit than our training results.

# Model 2

The second multiple linear regression model is as follows:

$$\text{BW} = \alpha + \beta_1 \text{MA} + \beta_2 \text{Gest} + \beta_3 \text{Nat} + \beta_4 \text{Age} + \epsilon$$

whereas BW is birthweight in grams, MA is maternal ancestry (Euro-descent, Afro-descent), Gest is gestational status (nulliparous or multiparous), Age is maternal age in years, and $\epsilon$ is the error term. Results are displayed in Table 3.

```{r}
# Final model 2 for paper using all data
# Using recipes and workflows
# Define recipe for model
ML_recipe2 <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") #%>% # Set reference level for ModifiedColor
  #step_dummy(ModifiedColor) #%>% # Convert all nominal variables to dummy variables. Cannot do this if I am merging tables in gtsummary
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors. Cannot do this if I am merging tables in gtsummary

# Define model specification
ML_spec2 <- linear_reg() %>% 
  set_engine("lm") %>%
  set_mode("regression")

# Create workflow
ML_wf2 <- workflow() %>%
  add_recipe(ML_recipe2) %>%
  add_model(ML_spec2)

# Fit the workflow
ML_fit2 <- fit(ML_wf2, data = ML_linear)

# Summarize model fit with tidiers
ML_lm2_summary <- summary(ML_fit2) # Summary of fit
tidy(ML_fit2, conf.int = TRUE) %>% mutate(p.value = round(p.value, 5)) # Coefficients and confidence intervals with p-values rounded to 5 decimal places
glance(ML_fit2) # Model-level statistics

# Produce predictions for R-squared
ML_fit2_pred <- predict(ML_fit2, new_data = ML_linear %>% select(-Weightgrams))
ML_fit2_pred

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit2_pred <- bind_cols(ML_fit2_pred, ML_linear %>% select(Weightgrams))
ML_fit2_pred

# Plot observed versus predicted
plot2 <- ggplot(ML_fit2_pred, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot2

# Create metric set including RMSE and R-squared
metrics_fit2 <- metric_set(rmse, rsq)
metrics_fit2(ML_fit2_pred, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit2 <- as.data.frame(metrics_fit2(ML_fit2_pred, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit2

# Create a gtsummary table directly from the model
table_model2 <- tbl_regression(ML_fit2, exponentiate = FALSE, intercept = TRUE) %>%
  add_significance_stars(hide_ci = TRUE, hide_p = TRUE) %>%
  add_glance_table(include = c(r.squared, nobs))
table_model2

# Save regression output into table
saveRDS(table_model2, file = here::here("results", "tables", "table_model2_final.rds"))
```
Our RMSE is 490 and our r-squared is 0.026.

## Model 2 Performance

Now let's run the model on our training data.

```{r}
# Training model 2
# Using recipes and workflows
# Define recipe for model
ML_recipe2_train <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = train_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Define model specification
ML_spec2_train <- linear_reg() %>% 
  set_engine("lm") %>%
  set_mode("regression")

# Create workflow
ML_wf2_train <- workflow() %>%
  add_recipe(ML_recipe2_train) %>%
  add_model(ML_spec2_train)

# Fit the workflow
ML_fit2_train <- fit(ML_wf2_train, data = train_data)

# Summarize model fit with tidiers
ML_lm2_summary_train <- summary(ML_fit2_train) # Summary of fit
tidy(ML_fit2_train, conf.int = TRUE) %>% mutate(p.value = round(p.value, 5)) # Coefficients and confidence intervals with p-values rounded to 5 decimal places
glance(ML_fit2_train) # Model-level statistics

# Produce predictions for R-squared
ML_fit2_pred_train <- predict(ML_fit2_train, new_data = train_data %>% select(-Weightgrams))
ML_fit2_pred_train

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit2_pred_train <- bind_cols(ML_fit2_pred_train, train_data %>% select(Weightgrams))
ML_fit2_pred_train

# Plot observed versus predicted
plot2_train <- ggplot(ML_fit2_pred_train, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot2_train

# Create metric set including RMSE and R-squared
metrics_fit2_train <- metric_set(rmse, rsq)
metrics_fit2_train(ML_fit2_pred_train, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit2_train <- as.data.frame(metrics_fit2_train(ML_fit2_pred_train, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit2_train

# Create a gtsummary table directly from the model
table_model2_train <- pull_workflow_fit(ML_fit2_train) %>%
  tbl_regression(exponentiate = FALSE, intercept = TRUE) %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = TRUE, hide_p = TRUE) %>%
  bold_p()
table_model2_train

# Save regression output into table
saveRDS(table_model2_train, file = here::here("results", "tables", "table_model2_train.rds"))
```
Our RMSE is 483 and our r-squared is 0.03, a slightly better performance than the full model.

Now let's run cross-validation.

```{r, message=FALSE,  warning=FALSE}
# CV model 2
# Set the seed for reproducibility
set.seed(rngseed)

# Create 10-fold cross-validation
folds <- vfold_cv(train_data, v = 10)
folds

# Define recipe for model
ML_recipe2_cv <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = train_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Set model specification
ML_spec2_cv <- linear_reg() %>% set_engine("lm")

# Create and initialize workflow that bundles together model specification and formula.
ML_wf2_cv <- workflow() %>% 
  add_model(ML_spec2_cv) %>% 
  add_recipe(ML_recipe2_cv)
  fit_resamples(ML_wf2_cv, resamples = folds)

# Resample the model
ML_wf2_cv_rs <- fit_resamples(ML_wf2_cv, resamples = folds)
ML_wf2_cv_rs

# Get model metrics
ML_cv2_metrics <- collect_metrics(ML_wf2_cv_rs)

# Filter the data frame to include only rows with "rmse" and "rsq" in the ".metric" column
ML_cv2_metrics <- ML_cv2_metrics[ML_cv2_metrics$.metric %in% c("rmse", "rsq"), !(names(ML_cv2_metrics) %in% ".config")]
ML_cv2_metrics

# Save regression output into table
saveRDS(ML_cv2_metrics, file = here::here("results", "tables", "ML_cv2_metrics.rds"))
```
Here, our RMSE is 483 and our r-squared is 0.026, the same as the training model.

Finally, let's run our model on the test data.

```{r}
# Testing model 2
# Using recipes and workflows
# Define recipe for model
ML_recipe2_test <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = test_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Define model specification
ML_spec2_test <- linear_reg() %>% 
  set_engine("lm") %>%
  set_mode("regression")

# Create workflow
ML_wf2_test <- workflow() %>%
  add_recipe(ML_recipe2_test) %>%
  add_model(ML_spec2_test)

# Fit the workflow
ML_fit2_test <- fit(ML_wf2_test, data = test_data)

# Summarize model fit with tidiers
ML_lm2_summary_test <- summary(ML_fit2_test) # Summary of fit
tidy(ML_fit2_test, conf.int = TRUE) %>% mutate(p.value = round(p.value, 5)) # Coefficients and confidence intervals with p-values rounded to 5 decimal places
glance(ML_fit2_test) # Model-level statistics

# Produce predictions for R-squared
ML_fit2_pred_test <- predict(ML_fit2_test, new_data = test_data %>% select(-Weightgrams))
ML_fit2_pred_test

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit2_pred_test <- bind_cols(ML_fit2_pred_test, test_data %>% select(Weightgrams))
ML_fit2_pred_test

# Plot observed versus predicted
plot2_test <- ggplot(ML_fit2_pred_test, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot2_test

# Create metric set including RMSE and R-squared
metrics_fit2_test <- metric_set(rmse, rsq)
metrics_fit2_test(ML_fit2_pred_test, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit2_test <- as.data.frame(metrics_fit2_test(ML_fit2_pred_test, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit2_test

# Create a gtsummary table directly from the model
table_model2_test <- pull_workflow_fit(ML_fit2_test) %>%
  tbl_regression(exponentiate = FALSE, intercept = TRUE) %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = TRUE, hide_p = TRUE) %>%
  bold_p()
table_model2_test

# Save regression output into table
saveRDS(table_model2_test, file = here::here("results", "tables", "table_model2_test.rds"))
```
Running the model on our test data, we see no significant results at any alpha level.

# Merging Model 1 and 2 into Table

```{r}
# Combine the two tables using tbl_merge
table3 <- tbl_merge(list(table_model1, table_model2), tab_spanner = c("**Model 1**", "**Model 2**")) %>%
  modify_table_body(fun = ~.x %>% arrange(variable))
table3

# Save regression output into table
saveRDS(table3, file = here::here("results", "tables", "table3_final.rds"))
```

# Model 3

The second two models are also simple and multiple linear regressions with outcome variable as birthweight in grams. However, in these two models, I maintained maternal skin color as three categories, White (the reference group), Black, and mixed race to see if there is differential outcomes for Black and mixed-race women. The first simple linear regression model is as follows:

$$\text{BW} = \alpha + \beta_1 \text{MC} + \epsilon$$
whereas BW is birthweight in grams, MC is maternal skin color (White, Black, and mixed race), and $\epsilon$ is the error term. Results are displayed in Table 4.

```{r}
# Final model 3 for paper using all data
# Using recipes and workflows
# Simple linear model
# Specify recipe
ML_recipe3 <- recipe(Weightgrams ~ Color, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White") #%>% # Set reference level for White
  #step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Specify model
ML_spec3 <- linear_reg() %>%
  set_engine("lm")

# Create workflow
ML_wf3 <- workflow() %>%
  add_recipe(ML_recipe3) %>%
  add_model(ML_spec3)

# Fit the workflow
ML_fit3 <- fit(ML_wf3, data = ML_linear)

# Summarize model fit with tidiers and create a gtsummary table directly from the model
ML_summary3 <- summary(ML_fit3)
tidy(ML_fit3, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit3) # Model-level statistics

# Produce predictions for R-squared
ML_fit3_pred <- predict(ML_fit3, new_data = ML_linear %>% select(-Weightgrams))
ML_fit3_pred

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit3_pred <- bind_cols(ML_fit3_pred, ML_linear %>% select(Weightgrams))
ML_fit3_pred

# Plot observed versus predicted
plot3 <- ggplot(ML_fit3_pred, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot3

# Create metric set including RMSE and R-squared
metrics_fit3 <- metric_set(rmse, rsq)
metrics_fit3(ML_fit3_pred, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit3 <- as.data.frame(metrics_fit3(ML_fit3_pred, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit3

# Create a gtsummary table from model
table_model3 <- tbl_regression(ML_fit3, exponentiate = FALSE, intercept = TRUE) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = TRUE, hide_p = TRUE) %>%
  add_glance_table(include = c(r.squared, nobs))
table_model3

# Save regression output into table
saveRDS(table_model3, file = here("results", "tables", "table_model3_final.rds"))
```

Our RMSE is 494 and our r-squared is 0.01.

## Model 3 Performance

Let's now evaluate the model on our training data.

```{r, message=FALSE,  warning=FALSE}
# Training model 3
# Using recipes and workflows
# Specify recipe
ML_recipe3_train <- recipe(Weightgrams ~ Color, data = train_data) %>%
  step_relevel(Color, ref_level = "White")

# Specify model
ML_spec3_train <- linear_reg() %>%
  set_engine("lm")

# Create workflow
ML_wf3_train <- workflow() %>%
  add_recipe(ML_recipe3_train) %>%
  add_model(ML_spec3_train)

# Fit the workflow
ML_fit3_train <- fit(ML_wf3_train, data = train_data)

# Summarize model fit with tidiers and create a gtsummary table directly from the model
ML_summary3_train <- summary(ML_fit3_train)
tidy(ML_fit3_train, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit3_train) # Model-level statistics

# Produce predictions for R-squared
ML_fit3_pred_train <- predict(ML_fit3_train, new_data = train_data %>% select(-Weightgrams))
ML_fit3_pred_train

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit3_pred_train <- bind_cols(ML_fit3_pred_train, train_data %>% select(Weightgrams))
ML_fit3_pred_train

# Plot observed versus predicted
plot3_train <- ggplot(ML_fit3_pred_train, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot3_train

# Create metric set including RMSE and R-squared
metrics_fit3_train <- metric_set(rmse, rsq)
metrics_fit3_train(ML_fit3_pred_train, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit3_train <- as.data.frame(metrics_fit3_train(ML_fit3_pred_train, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit3_train

# Create a gtsummary table from model
table_model3_train <- tbl_regression(ML_fit3_train, exponentiate = FALSE, intercept = TRUE) %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = TRUE, hide_p = TRUE)
table_model3_train

# Save regression output into table
saveRDS(table_model3_train, file = here("results", "tables", "table_model3_train.rds"))
```

Our RMSE is 489 and our r-squared is 0.01, slightly better performance than the full model. Our results are not significantly different from the full model.

Now let's run cross-validation.

```{r, message=FALSE,  warning=FALSE}
# CV model 3
# Set the seed for reproducibility
set.seed(rngseed)

# Create 10-fold cross-validation
folds <- vfold_cv(train_data, v = 10)
folds

# Define recipe for model
ML_recipe3_cv <- recipe(Weightgrams ~ Color, data = train_data) %>%
  step_relevel(Color, ref_level = "White")

# Set model specification
ML_spec3_cv <- linear_reg() %>% set_engine("lm")

# Create and initialize workflow that bundles together model specification and formula.
ML_wf3_cv <- workflow() %>% 
  add_model(ML_spec3_cv) %>% 
  add_recipe(ML_recipe3_cv)
  fit_resamples(ML_wf3_cv, resamples = folds)

# Resample the model
ML_wf3_cv_rs <- fit_resamples(ML_wf3_cv, resamples = folds)
ML_wf3_cv_rs

# Get model metrics
collect_metrics(ML_wf3_cv_rs)

# Get model metrics
ML_cv3_metrics <- collect_metrics(ML_wf3_cv_rs)

# Filter the data frame to include only rows with "rmse" and "rsq" in the ".metric" column
ML_cv3_metrics <- ML_cv3_metrics[ML_cv3_metrics$.metric %in% c("rmse", "rsq"), !(names(ML_cv3_metrics) %in% ".config")]
ML_cv3_metrics

# Save regression output into table
saveRDS(ML_cv3_metrics, file = here::here("results", "tables", "ML_cv3_metrics.rds"))
```
Using cross-validation, our RMSE is 487 and our r-squared is 0.01, a similar performance to our training model.

Finally, let's run our model on the test data.

```{r, message=FALSE, warning=FALSE}
# Testing model 3
# Using recipes and workflows
# Specify recipe
ML_recipe3_test <- recipe(Weightgrams ~ Color, data = test_data) %>%
  step_relevel(Color, ref_level = "White")

# Specify model
ML_spec3_test <- linear_reg() %>%
  set_engine("lm")

# Create workflow
ML_wf3_test <- workflow() %>%
  add_recipe(ML_recipe3_test) %>%
  add_model(ML_spec3_test)

# Fit the workflow
ML_fit3_test <- fit(ML_wf3_test, data = test_data)

# Summarize model fit with tidiers and create a gtsummary table directly from the model
ML_summary3_test <- summary(ML_fit3_test)
tidy(ML_fit3_test, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit3_test) # Model-level statistics

# Produce predictions for R-squared
ML_fit3_pred_test <- predict(ML_fit3_test, new_data = test_data %>% select(-Weightgrams))
ML_fit3_pred_test

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit3_pred_test <- bind_cols(ML_fit3_pred_test, test_data %>% select(Weightgrams))
ML_fit3_pred_test

# Plot observed versus predicted
plot3_test <- ggplot(ML_fit3_pred_test, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot3_test

# Create metric set including RMSE and R-squared
metrics_fit3_test <- metric_set(rmse, rsq)
metrics_fit3_test(ML_fit3_pred_test, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit3_test <- as.data.frame(metrics_fit3_test(ML_fit3_pred_test, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit3_test

# Create a gtsummary table from model
table_model3_test <- tbl_regression(ML_fit3_test, exponentiate = FALSE, intercept = TRUE) %>%
  bold_p() %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE)
table_model3_test

# Save regression output into table
saveRDS(table_model3_test, file = here("results", "tables", "table_model3_test.rds"))
```

Much like in the other testing models, our results are not significant (except the intercept). Our performance is also worse (RMSE of 507 and r-squared of 0.008).

# Model 4

The second multiple linear regression model is as follows:

$$\text{BW} = \alpha + \beta_1 \text{MC} + \beta_2 \text{Gest} + \beta_3 \text{Nat} + \beta_4 \text{Age} + \epsilon$$

whereas BW is birthweight in grams, MC is maternal skin color (White, Black and mixed race), Gest is gestational status (nullipara, primipara, multipara, multigravida), Nat is maternal nationality (Brazilian, Latin American, European, Middle Eastern), Age is maternal age in years, and $\epsilon$ is the error term.

```{r}
# Final model 4 for paper using all data
# Using recipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4 <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") #%>% # Set reference level for ModifiedStatus
  #step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Specify model
ML_spec4 <- linear_reg() %>%
  set_engine("lm")

# Create workflow
ML_wf4 <- workflow() %>%
  add_recipe(ML_recipe4) %>%
  add_model(ML_spec4)

# Fit the workflow
ML_fit4 <- fit(ML_wf4, data = ML_linear)

# Summarize model fit with tidiers and create a gtsummary table directly from the model
ML_summary4 <- summary(ML_fit4)
tidy(ML_fit4, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit4) # Model-level statistics

# Produce predictions for R-squared
ML_fit4_pred <- predict(ML_fit4, new_data = ML_linear %>% select(-Weightgrams))
ML_fit4_pred

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit4_pred <- bind_cols(ML_fit4_pred, ML_linear %>% select(Weightgrams))
ML_fit4_pred

# Plot observed versus predicted
plot4 <- ggplot(ML_fit4_pred, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot4

# Create metric set including RMSE and R-squared
metrics_fit4 <- metric_set(rmse, rsq)
metrics_fit4(ML_fit4_pred, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit4 <- as.data.frame(metrics_fit4(ML_fit4_pred, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit4

# Create a gtsummary table from model
table_model4 <- tbl_regression(ML_fit4, exponentiate = FALSE, intercept = TRUE) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = TRUE, hide_p = TRUE) %>% # Hide confidence intervals and p-values
    add_glance_table(include = c(r.squared, nobs)) # Add R-squared and number of observations
table_model4

# Save regression output into table
saveRDS(table_model4, file = here("results", "tables", "table_model4_final.rds"))
```
Our RMSE is 490 and our r-squared is 0.026.

## Model 4 Performance

Let's now evaluate the model on our training data.

```{r, message=FALSE,  warning=FALSE}
# Model 4 training
# Using recipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4_train <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = train_data) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") %>% # Set reference level for ModifiedStatus
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Specify model
ML_spec4_train <- linear_reg() %>%
  set_engine("lm")

# Create workflow
ML_wf4_train <- workflow() %>%
  add_recipe(ML_recipe4_train) %>%
  add_model(ML_spec4_train)

# Fit the workflow
ML_fit4_train <- fit(ML_wf4_train, data = train_data)

# Summarize model fit with tidiers and create a gtsummary table directly from the model
ML_summary4_train <- summary(ML_fit4_train)
tidy(ML_fit4_train, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit4_train) # Model-level statistics

# Produce predictions for R-squared
ML_fit4_pred_train <- predict(ML_fit4_train, new_data = train_data %>% select(-Weightgrams))
ML_fit4_pred_train

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit4_pred_train <- bind_cols(ML_fit4_pred_train, train_data %>% select(Weightgrams))
ML_fit4_pred_train

# Plot observed versus predicted
plot4_train <- ggplot(ML_fit4_pred_train, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot4_train

# Create metric set including RMSE and R-squared
metrics_fit4_train <- metric_set(rmse, rsq)
metrics_fit4_train(ML_fit4_pred_train, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit4_train <- as.data.frame(metrics_fit4_train(ML_fit4_pred_train, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit4_train

# Create a gtsummary table from model
table_model4_train <- tbl_regression(ML_fit4_train, exponentiate = FALSE, intercept = TRUE) %>%
  bold_p() %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE)
table_model4_train

# Save regression output into table
saveRDS(table_model4_train, file = here("results", "tables", "table_model4_train.rds"))
```
Our RMSE is 483 and our r-squared is 0.033, a slightly better performance than the full model.

Now let's run cross-validation.

```{r, message=FALSE,  warning=FALSE}
# CV model 4
# Set the seed for reproducibility
set.seed(rngseed)

# Create 10-fold cross-validation
folds <- vfold_cv(train_data, v = 10)
folds

# Define recipe for model
ML_recipe4_cv <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") %>% # Set reference level for ModifiedStatus
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Set model specification
ML_spec4_cv <- linear_reg() %>% set_engine("lm")

# Create and initialize workflow that bundles together model specification and formula.
ML_wf4_cv <- workflow() %>% 
  add_model(ML_spec4_cv) %>% 
  add_recipe(ML_recipe4_cv)
  fit_resamples(ML_wf4_cv, resamples = folds)

# Resample the model
ML_wf4_cv_rs <- fit_resamples(ML_wf4_cv, resamples = folds)
ML_wf4_cv_rs

# Get model metrics
collect_metrics(ML_wf4_cv_rs)

# Get model metrics
ML_cv4_metrics <- collect_metrics(ML_wf4_cv_rs)

# Filter the data frame to include only rows with "rmse" and "rsq" in the ".metric" column
ML_cv4_metrics <- ML_cv4_metrics[ML_cv4_metrics$.metric %in% c("rmse", "rsq"), !(names(ML_cv4_metrics) %in% ".config")]
ML_cv4_metrics

# Save regression output into table
saveRDS(ML_cv4_metrics, file = here::here("results", "tables", "ML_cv4_metrics.rds"))
```
Here, our RMSE is 484 and our r-squared is 0.025, a better performance than both the full model and training model.

Finally, let's run our model on the test data.

```{r, message=FALSE, warning=FALSE}
# Model 4 testing
# Using recipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4_test <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = test_data) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") %>% # Set reference level for ModifiedStatus
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Specify model
ML_spec4_test <- linear_reg() %>%
  set_engine("lm")

# Create workflow
ML_wf4_test <- workflow() %>%
  add_recipe(ML_recipe4_test) %>%
  add_model(ML_spec4_test)

# Fit the workflow
ML_fit4_test <- fit(ML_wf4_test, data = test_data)

# Summarize model fit with tidiers and create a gtsummary table directly from the model
ML_summary4_test <- summary(ML_fit4_test)
tidy(ML_fit4_test, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit4_test) # Model-level statistics

# Produce predictions for R-squared
ML_fit4_pred_test <- predict(ML_fit4_test, new_data = test_data %>% select(-Weightgrams))
ML_fit4_pred_test

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit4_pred_test <- bind_cols(ML_fit4_pred_test, test_data %>% select(Weightgrams))
ML_fit4_pred_test

# Plot observed versus predicted
plot4_test <- ggplot(ML_fit4_pred_test, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot4_test

# Create metric set including RMSE and R-squared
metrics_fit4_test <- metric_set(rmse, rsq)
metrics_fit4_test(ML_fit4_pred_test, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit4_test <- as.data.frame(metrics_fit4_test(ML_fit4_pred_test, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit4_test

# Create a gtsummary table from model
table_model4_test <- tbl_regression(ML_fit4_test, exponentiate = FALSE, intercept = TRUE) %>%
  bold_p() %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE)
table_model4_test

# Save regression output into table
saveRDS(table_model4_test, file = here("results", "tables", "table_model4_test.rds"))
```
Again, most of our results are not significant except the intercept given the small sample size.

## Merge Models 3 and 4
```{r}
# Combine the two tables using tbl_merge
table4 <- tbl_merge(list(table_model3, table_model4), tab_spanner = c("**Model 3**", "**Model 4**")) %>%
  modify_table_body(fun = ~.x %>% arrange(variable))
table4

# Save regression output into table
saveRDS(table4, file = here("results", "tables", "table4_final.rds"))
```