### 2024-06-29 code/analysis-code/introanalysis-v1.qmd

---
title: "Intro Statistical Analysis Script"
author: "Cassia Roth"
date: "`r Sys.Date()`"
output: html_document
---

# Setup

```{r, message=FALSE,  warning=FALSE}
#load needed packages. make sure they are installed.
library(tidyverse) #for data processing/cleaning; includes ggplot2, tidyr, readr, dplyr, stringr, purr, forcats
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(renv) #for package management
library(knitr) #for nice tables
library(kableExtra) #for nice tables
library(gt) #for nice tables
library(gtsummary) #for summary tables
library(ggplot2) #for plotting
library(tidymodels) #for modeling
rm(list=ls()) #clear the environment
```

Load the data

```{r}
#Path to summary data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","ML_linear.rds")
#load data
ML_linear <- readRDS(data_location)
```

## Linear Models

Let's run a simple regression with the outcome variable (birth weight in grams) and the exposure variable maternal ancestry. I will use the recipe and workflow features of `tidymodels`.

```{r}
# Model using recipes and workflows
# Define recipe for model
ML_recipe1 <- recipe(Weightgrams ~ ModifiedColor, data = ML_linear) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable

# Specify model
ML_spec1 <- linear_reg() %>% 
  set_engine("lm")

# Create work flow
ML_wf1 <- workflow() %>%
  add_recipe(ML_recipe1) %>%
  add_model(ML_spec1)

# Fit the work flow
ML_fit1 <- fit(ML_wf1, data = ML_linear)

# Summarize model fit with tidiers
ML_summary1 <- summary(ML_fit1) # Summary of fit
tidy(ML_fit1, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit1) # Model-level statistics

# Produce predictions for R-squared
ML_fit1_pred <- predict(ML_fit1, new_data = ML_linear %>% select(-Weightgrams))
ML_fit1_pred

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit1_pred <- bind_cols(ML_fit1_pred, ML_linear %>% select(Weightgrams))
ML_fit1_pred

# Plot observed versus predicted
plot1 <- ggplot(ML_fit1_pred, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot1

# Create metric set including RMSE and R-squared
metrics_fit1 <- metric_set(rmse, rsq)
metrics_fit1(ML_fit1_pred, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit1 <- as.data.frame(metrics_fit1(ML_fit1_pred, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit1

# Create a gtsummary table directly from the model
table_fit1 <-
  tbl_regression(ML_fit1, exponentiate = FALSE, intercept = TRUE, show_single_row = everything()) %>% 
  bold_p() %>% # make p-values bold
  add_glance_table(include = c(r.squared, adj.r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error))

#Print table
table_fit1

# Save regression output into table
saveRDS(table_fit1, file = here::here("results", "tables", "table_fit1.rds"))
```
Now let's run a second simple bivariate analysis, looking at infant birth weight as a function of maternal age. This time, when plotting weight as a function of maternal age, we can use a scatterplot.

```{r}
# Model using recipes and workflows
# Create recipe
ML_recipe_age <- recipe(Weightgrams ~ Age, data = ML_linear)

# Specify model
ML_spec_age <- linear_reg() %>% 
  set_engine("lm")

# Create work flow
ML_wf_age <- workflow() %>%
  add_recipe(ML_recipe_age) %>%
  add_model(ML_spec_age)

# Fit the work flow
ML_fit_age <- fit(ML_wf_age, data = ML_linear)

# Summarize model fit with tidiers
ML_summary_age <- summary(ML_fit_age) # Summary of fit
tidy(ML_fit_age, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit_age) # Model-level statistics

# Produce predictions for RMSE and R-squared
ML_age_pred <- predict(ML_fit_age, new_data = ML_linear %>% select(-Weightgrams))
ML_age_pred

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_age_pred <- bind_cols(ML_age_pred, ML_linear %>% select(Weightgrams))
ML_age_pred

# Plot observed versus predicted
plot_age <- ggplot(ML_age_pred, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot_age

# Create metric set including RMSE and R-squared
metrics_fit_age <- metric_set(rmse, rsq)
metrics_fit_age(ML_age_pred, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit_age <- as.data.frame(metrics_fit_age(ML_age_pred, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit_age

# Create a gtsummary table directly from the model
table_fit2 <-
  tbl_regression(ML_fit_age, exponentiate = FALSE, intercept = TRUE, show_single_row = everything()) %>% 
  bold_p() %>% 
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error))

#Print table
table_fit2

# Save regression output into table
saveRDS(table_fit2, file = here::here("results", "tables", "table_fit2.rds"))
```

Now let's run a slightly more complicated model that includes maternal skin color, age, nationality, and gestational status.

```{r}
# Set reference level for ModifiedColor
ML_linear$ModifiedColor <- relevel(ML_linear$ModifiedColor, ref = "Euro-Descent")

# Specify and fit model using linear_reg() (default engine, OLS) from tidymodels
ML_lm3 <- linear_reg() %>% set_engine("lm") %>% fit(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = ML_linear)

# Summarize model fit with tidiers
ML_lm3_summary <- summary(ML_lm3) # Summary of fit
tidy(ML_lm3, conf.int = TRUE) %>% mutate(p.value = round(p.value, 5)) # Coefficients and confidence intervals with p-values rounded to 5 decimal places
glance(ML_lm3) # Model-level statistics

# Create a gtsummary table directly from the model
table_fit3 <- tbl_regression(ML_lm3, exponentiate = FALSE, intercept = TRUE) %>% 
  bold_p() %>% 
  add_glance_table(include = c(r.squared, adj.r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error))
table_fit3

# Save regression output into table
saveRDS(table_fit3, file = here::here("results", "tables", "table_fit3.rds"))
```

In the first and third models, there appears to be an association between maternal skin color and infant birth weight, with Euro-descended women (the reference group) having infants with higher birth weights than Afro-descended women. In the second model, we can see that older mothers are associated with giving birth to infants with higher birth weights.

For all three models, the r-squared is very small (0.009, 0.008, and 0.026), indicating that the model does not explain much of the variance in birth weight. This is likely due to the fact that birth weight is a complex trait influenced by many factors, including genetic, environmental, and social factors.

I am not trying to explain which general factors are trying to explain birth weight. Rather, I'm trying to understand if maternal skin color is associated with lower infant birth weight during this specific time period in Rio de Janeiro, Brazil. For example, the gestational age probably explains much of the variation in the model, yet this variable was only included in very few observations in the published clinic data from which I created this dataset.

Given my hypothesis, that the legacies of slavery affected maternal-fetal health, then maternal skin color, however, is also probably associated with premature birth, which is correlated to infant birth weight.

Now let's try to run a model where skin color is differentiated according to the original data, or with three categories (White, Black, and Mixed Race). We'll start with a simple model of infant birth weight as a function of maternal skin color. Then we will run a model with all the predictors.

```{r}
# Simple linear model
# Set reference level for Color
ML_linear$Color <- relevel(ML_linear$Color, ref = "White")

# Specify and fit model using linear_reg() (default engine, OLS) from tidymodels
ML_lm4 <- linear_reg() %>% set_engine("lm") %>% fit(Weightgrams ~ Color, data = ML_linear)

# Summarize model fit with tidiers
ML_lm4_summary <- summary(ML_lm4) # Summary of fit
tidy(ML_lm4, conf.int = TRUE) %>% mutate(p.value = round(p.value, 5)) # Coefficients and confidence intervals with p-values rounded to 5 decimal places
glance(ML_lm4) # Model-level statistics

# Create a gtsummary table directly from the model
table_fit4 <- tbl_regression(ML_lm4, exponentiate = FALSE, intercept = TRUE) %>% 
  bold_p() %>% 
  add_glance_table(include = c(r.squared, adj.r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error))
table_fit4

# Save regression output into table
saveRDS(table_fit4, file = here::here("results", "tables", "table_fit4.rds"))

# Complex linear model
# Set reference level for Color
ML_linear$Color <- relevel(ML_linear$Color, ref = "White")

# Specify and fit model using linear_reg() (default engine, OLS) from tidymodels
ML_lm4 <- linear_reg() %>% set_engine("lm") %>% fit(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = ML_linear)

# Summarize model fit with tidiers
ML_lm4_summary <- summary(ML_lm4) # Summary of fit
tidy(ML_lm4, conf.int = TRUE) %>% mutate(p.value = round(p.value, 5)) # Coefficients and confidence intervals with p-values rounded to 5 decimal places
glance(ML_lm4) # Model-level statistics

# Create a gtsummary table directly from the model
table_fit4 <- tbl_regression(ML_lm4, exponentiate = FALSE, intercept = TRUE) %>% 
  bold_p() %>% 
  add_glance_table(include = c(r.squared, adj.r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error))
table_fit4

# Save regression output into table
saveRDS(table_fit4, file = here::here("results", "tables", "table_fit4.rds"))
```

## Logistic models

Now we will still consider `Birthweight` as the outcome of interest, but we will transform it into a categorical outcome of <=2500 (low birthweight, LBW) and >2500 (normal birthweight, NBW) and fit a logistic model, using the main predictor of interest, `ModifiedColor`.

First, we will split data into training and testing sets to later calculate ROC_AUC.

```{r}
set.seed(123)
ML_log_split <- initial_split(ML_linear, prop = 0.75, strata = BirthweightCategory)

# Create training and test datasets
ML_log_train <- training(ML_log_split)
ML_log_test <- testing(ML_log_split)
```

Now we will run the first logistic model with `ModifiedColor` as the predictor.

```{r}
# Simple logistic regression model with ModifiedColor as the predictor
# Check factor levels
levels(ML_linear$BirthweightCategory) # Reference is LBW

# Set the reference level explicitly to NBW
ML_linear <- ML_linear %>%
  mutate(BirthweightCategory = relevel(BirthweightCategory, ref = "NBW"))

# Create recipe
# Create recipe simple logistic model (creating dummy variables and standardizing continuous variables)
log1_recipe <- recipe(BirthweightCategory ~ ModifiedColor, data = ML_linear) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_normalize(all_predictors())

# Define model specification
log1_spec <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")

# Create workflow
log1_wf <- workflow() %>%
  add_recipe(log1_recipe) %>%
  add_model(log1_spec)

# Fit model
log1_fit <- log1_wf %>%
fit(data = ML_linear)

# Extract fitted model from workflow
log1_fitted <- extract_fit_parsnip(log1_fit)$fit

# Create a gtsummary table directly from the model
table_log1 <- tbl_regression(log1_fitted, exponentiate = TRUE) # use exponentiate true for ORs
table_log1

# Save regression output into table
saveRDS(table_log1, file = here::here("results", "tables", "table_log1.rds"))

# Calculate performance metrics using pseudo-R2
# Extract model metrics using broom
model_metrics1 <- glance(log1_fitted)

# Calculate McFadden's R-squared
# McFadden's R-squared = 1 - (deviance(model) / deviance(null model))
mcfadden_r2_1 <- 1 - (model_metrics1$deviance / model_metrics1$null.deviance)

# Print McFadden's R-squared
print(mcfadden_r2_1)

# Calculate performance metrics using ROC_AUC
# Fit the model on the training data
log1_fit <- log1_wf %>% fit(data = ML_log_train)

# Predict probabilities on the test set
prob_preds1 <- predict(log1_fit, new_data = ML_log_test, type = "prob")

# Combine predictions with true values
results1 <- bind_cols(ML_log_test, prob_preds1)

# Calculate ROC-AUC
roc_auc_result1 <- roc_auc(results1, truth = BirthweightCategory, .pred_LBW)

# Print ROC-AUC
print(roc_auc_result1)

# Create and plot the ROC curve
roc_curve_data1 <- roc_curve(results1, truth = BirthweightCategory, .pred_LBW)

# Plotting the ROC curve
roc_plot1 <- ggplot(roc_curve_data1, aes(x = 1 - specificity, y = sensitivity)) +
  geom_path() +
  geom_abline(lty = 2) +
  labs(title = "ROC Curve",
       x = "1 - Specificity",
       y = "Sensitivity")

# Print the plot
print(roc_plot1)
```
With an ROC_AUC of 0.463, our model is performing worse than random guessing. Our pseudo R-squared is 0.004.

Now let's run another logistic model with `Color` as the predictor.

```{r}
# Simple logistic regression model with ModifiedColor as the predictor
# Check factor levels
levels(ML_linear$BirthweightCategory) # Reference is LBW

# Set the reference level explicitly to NBW
ML_linear <- ML_linear %>%
  mutate(BirthweightCategory = relevel(BirthweightCategory, ref = "NBW"))

# Create recipe
# Create recipe simple logistic model (creating dummy variables and standardizing continuous variables)
log2_recipe <- recipe(BirthweightCategory ~ Color, data = ML_linear) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_normalize(all_predictors())

# Define model specification
log2_spec <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")

# Create workflow
log2_wf <- workflow() %>%
  add_recipe(log2_recipe) %>%
  add_model(log2_spec)

# Fit model
log2_fit <- log2_wf %>%
fit(data = ML_linear)

# Extract fitted model from workflow
log2_fitted <- extract_fit_parsnip(log2_fit)$fit

# Create a gtsummary table directly from the model
table_log2 <- tbl_regression(log2_fitted, exponentiate = TRUE) # use exponentiate true for ORs
table_log2

# Save regression output into table
saveRDS(table_log2, file = here::here("results", "tables", "table_log2.rds"))

# Calculate metrics using pseudo-R2
# Extract model metrics using broom
model_metrics2 <- glance(log2_fitted)

# Calculate McFadden's R-squared
# McFadden's R-squared = 1 - (deviance(model) / deviance(null model))
mcfadden_r2_2 <- 1 - (model_metrics2$deviance / model_metrics2$null.deviance)

# Print McFadden's R-squared
print(mcfadden_r2_2)

# Calculate performance metrics using ROC_AUC
# Fit the model on the training data
log2_fit <- log2_wf %>% fit(data = ML_log_train)

# Predict probabilities on the test set
prob_preds2 <- predict(log2_fit, new_data = ML_log_test, type = "prob")

# Combine predictions with true values
results2 <- bind_cols(ML_log_test, prob_preds2)

# Calculate ROC-AUC
roc_auc_result2 <- roc_auc(results2, truth = BirthweightCategory, .pred_LBW)

# Print ROC-AUC
print(roc_auc_result2)

# Create and plot the ROC curve
roc_curve_data2 <- roc_curve(results2, truth = BirthweightCategory, .pred_LBW)

# Plotting the ROC curve
roc_plot2 <- ggplot(roc_curve_data2, aes(x = 1 - specificity, y = sensitivity)) +
  geom_path() +
  geom_abline(lty = 2) +
  labs(title = "ROC Curve",
       x = "1 - Specificity",
       y = "Sensitivity")

# Print the plot
print(roc_plot2)
```
The ROC_AUC for our second simple logistic regression model is 0.47, or still worse than random guessing. Our pseudo R-squared is 0.006.

Now, we will run two more logistic models, one with `ModifiedColor` and the other covariates `Age`, `ModifiedStatus`, `ModifiedNationality`, and one with `Color` and these same covariates.

```{r}
# Multiple logistic regression model with ModifiedColor as the predictor
# Check factor levels
levels(ML_linear$BirthweightCategory) # Reference is LBW
levels(ML_linear$ModifiedColor) # Reference is Euro-descent
levels(ML_linear$ModifiedStatus) # Reference is multiparous
levels(ML_linear$ModifiedNationality) # Reference is European

# Set the reference level explicitly to NBW
ML_linear <- ML_linear %>%
  mutate(BirthweightCategory = relevel(BirthweightCategory, ref = "NBW"),
         ModifiedNationality = relevel(ModifiedNationality, ref = "Brazilian"))

# Create recipe
# Create recipe simple logistic model (creating dummy variables and standardizing continuous variables)
log3_recipe <- recipe(BirthweightCategory ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_normalize(all_predictors())

# Define model specification
log3_spec <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")

# Create workflow
log3_wf <- workflow() %>%
  add_recipe(log3_recipe) %>%
  add_model(log3_spec)

# Fit model
log3_fit <- log3_wf %>%
fit(data = ML_linear)

# Extract fitted model from workflow
log3_fitted <- extract_fit_parsnip(log3_fit)$fit

# Create a gtsummary table directly from the model
table_log3 <- tbl_regression(log3_fitted, exponentiate = TRUE) # use exponentiate true for ORs
table_log3 <- bold_p(table_log3, t = 0.05)
table_log3

# Save regression output into table
saveRDS(table_log3, file = here::here("results", "tables", "table_log3.rds"))

# Calculate metrics using pseudo-R2
# Extract model metrics using broom
model_metrics3 <- glance(log3_fitted)

# Calculate McFadden's R-squared
# McFadden's R-squared = 1 - (deviance(model) / deviance(null model))
mcfadden_r2_3 <- 1 - (model_metrics3$deviance / model_metrics3$null.deviance)

# Print McFadden's R-squared
print(mcfadden_r2_3)

# Calculate performance metrics using ROC_AUC
# Fit the model on the training data
log3_fit <- log3_wf %>% fit(data = ML_log_train)

# Predict probabilities on the test set
prob_preds3 <- predict(log3_fit, new_data = ML_log_test, type = "prob")

# Combine predictions with true values
results3 <- bind_cols(ML_log_test, prob_preds3)

# Calculate ROC-AUC
roc_auc_result3 <- roc_auc(results3, truth = BirthweightCategory, .pred_LBW)

# Print ROC-AUC
print(roc_auc_result3)

# Create and plot the ROC curve
roc_curve_data3 <- roc_curve(results3, truth = BirthweightCategory, .pred_LBW)

# Plotting the ROC curve
roc_plot3 <- ggplot(roc_curve_data3, aes(x = 1 - specificity, y = sensitivity)) +
  geom_path() +
  geom_abline(lty = 2) +
  labs(title = "ROC Curve",
       x = "1 - Specificity",
       y = "Sensitivity")

# Print the plot
print(roc_plot3)
```
Our ROC_AUC here is worse, at 0.46, or worse than random guessing. Our pseudo-R2 is 0.0055.

```{r}
# Multiple logistic regression model with Color as the predictor
# Check factor levels
levels(ML_linear$BirthweightCategory) # Reference is LBW
levels(ML_linear$Color) # Reference is White
levels(ML_linear$ModifiedStatus) # Reference is multiparous
levels(ML_linear$ModifiedNationality) # Reference is European

# Set the reference level explicitly to NBW
ML_linear <- ML_linear %>%
  mutate(BirthweightCategory = relevel(BirthweightCategory, ref = "NBW"),
         ModifiedNationality = relevel(ModifiedNationality, ref = "Brazilian"))

# Create recipe
# Create recipe simple logistic model (creating dummy variables and standardizing continuous variables)
log4_recipe <- recipe(BirthweightCategory ~ Color + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_normalize(all_predictors())

# Define model specification
log4_spec <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")

# Create workflow
log4_wf <- workflow() %>%
  add_recipe(log4_recipe) %>%
  add_model(log4_spec)

# Fit model
log4_fit <- log4_wf %>%
fit(data = ML_linear)

# Extract fitted model from workflow
log4_fitted <- extract_fit_parsnip(log4_fit)$fit

# Create a gtsummary table directly from the model
table_log4 <- tbl_regression(log4_fitted, exponentiate = TRUE) # exponentiate true for ORs

#Bold significant p-values
table_log4 <- bold_p(table_log4, t = 0.05)
table_log4

# Save regression output into table
saveRDS(table_log4, file = here::here("results", "tables", "table_log4.rds"))

# Calculate metrics using pseudo-R2
# Extract model metrics using broom
model_metrics4 <- glance(log4_fitted)

# Calculate McFadden's R-squared
# McFadden's R-squared = 1 - (deviance(model) / deviance(null model))
mcfadden_r2_4 <- 1 - (model_metrics4$deviance / model_metrics4$null.deviance)

# Print McFadden's R-squared
print(mcfadden_r2_4)

# Calculate performance metrics using ROC_AUC
# Fit the model on the training data
log4_fit <- log4_wf %>% fit(data = ML_log_train)

# Predict probabilities on the test set
prob_preds4 <- predict(log4_fit, new_data = ML_log_test, type = "prob")

# Combine predictions with true values
results4 <- bind_cols(ML_log_test, prob_preds4)

# Calculate ROC-AUC
roc_auc_result4 <- roc_auc(results4, truth = BirthweightCategory, .pred_LBW)

# Print ROC-AUC
print(roc_auc_result4)

# Create and plot the ROC curve
roc_curve_data4 <- roc_curve(results4, truth = BirthweightCategory, .pred_LBW)

# Plotting the ROC curve
roc_plot4 <- ggplot(roc_curve_data4, aes(x = 1 - specificity, y = sensitivity)) +
  geom_path() +
  geom_abline(lty = 2) +
  labs(title = "ROC Curve",
       x = "1 - Specificity",
       y = "Sensitivity")

# Print the plot
print(roc_plot4)
```
In our final model, our ROC_AUC is 0.477, still worse than random guessing. Our pseudo-R2 is 0.007.