### 2024-09-06 code/analysis-code/fullanalysis-v1.qmd
Final Revisions for Submission
Cleaned up all files

---
title: "Full Analysis Script"
author: "Cassia Roth"
date: "`r Sys.Date()`"
output: html_document
---

# Setup

```{r, message=FALSE,  warning=FALSE}
#load needed packages. make sure they are installed.
#install.packages("glue")
library(tidyverse) #for data processing/cleaning; includes ggplot2, tidyr, readr, dplyr, stringr, purr, forcats
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(renv) #for package management
library(knitr) #for nice tables
library(kableExtra) #for nice tables
library(gt) #for nice tables
library(gtsummary) #for summary tables
library(ggplot2) #for plotting
library(tidymodels) #for modeling
library(glue)
rm(list=ls()) #clear the environment
```

Load the data

```{r, message=FALSE,  warning=FALSE}
# Path to summary data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","ML_linear.rds")

# Load data
ML_linear <- readRDS(data_location)
str(ML_linear)
```

# Preparation for Model Assessment

First, I will create train and test data (75/25) to be able to evaluate the performance of the models. I will also set the seed to ensure reproducibility. For each model, I perform the following steps: 1) run the model on the train data; 2) perform cross-validation on the train data; 3) evaluate the model on the test data. The model outputs included in the final manuscript are on all the data.

```{r, message=FALSE,  warning=FALSE}
# Fix the random numbers by setting the seed 
# This enables the analysis to be reproducible when random numbers are used 
rngseed = 1234
set.seed(rngseed)

# Create training data
# Put 3/4 of the data into the training set 
data_split <- initial_split(ML_linear, prop = 3/4)

# Create data frames for the two sets:
train_data <- training(data_split)
test_data  <- testing(data_split)
```

# Model 1

The first simple linear regression model is as follows:

$$\text{BW} = \alpha + \beta_1 \text{MA} + \epsilon$$

whereas `BW` is birth weight in grams, $\alpha$ is the average birth weight, $\beta_1$ is the group deviation from the average birth weight, `MA` is maternal ancestry (Euro-descent, Afro-descent), and $\epsilon$ is the error term.

Full results are displayed in Table 2.

```{r, message=FALSE,  warning=FALSE}
# Model 1: Final model for paper using all data
# Create a list to store R-squared values
r2_1 <- list()
rmse_1 <- list()

# Model using recipes and workflows
# Define recipe for model
ML_recipe1 <- recipe(Weightgrams ~ ModifiedColor, data = ML_linear) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") #%>% # Set reference level for ModifiedColor
  #step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable. Cannot do this if I am merging tables

# Specify model
ML_spec1 <- linear_reg() %>% 
  set_engine("lm")

# Create work flow
ML_wf1 <- workflow() %>%
  add_recipe(ML_recipe1) %>%
  add_model(ML_spec1)

# Fit the work flow
ML_fit1 <- fit(ML_wf1, data = ML_linear)

# Summarize model fit with tidiers
ML_summary1 <- summary(ML_fit1) # Summary of fit
tidy(ML_fit1, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit1) # Model-level statistics

# Produce predictions for R-squared
ML_fit1_pred <- predict(ML_fit1, new_data = ML_linear %>% select(-Weightgrams))
ML_fit1_pred

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit1_pred <- bind_cols(ML_fit1_pred, ML_linear %>% select(Weightgrams))
ML_fit1_pred

# Plot observed versus predicted
plot1 <- ggplot(ML_fit1_pred, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot1

# Create metric set including RMSE and R-squared
metrics_fit1 <- metric_set(rmse, rsq)
metrics_fit1(ML_fit1_pred, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit1 <- as.data.frame(metrics_fit1(ML_fit1_pred, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit1

# Create a gtsummary table directly from the model
table_model1 <-
  tbl_regression(ML_fit1, exponentiate = FALSE, intercept = TRUE) %>% 
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = TRUE, hide_p = TRUE) %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_header(statistic = "**T-Statistic**") # rename header from statistic to t-statistic
table_model1

# Print table
table_model1

# Extract RMSE and R2 for in-line text
r2_1$model1 <- metrics_fit1[metrics_fit1$.metric == "rsq", ".estimate"]
rmse_1$model1 <- metrics_fit1[metrics_fit1$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model1, file = here::here("results", "tables", "table_model1_final.rds"))

# Create a gtsummary table directly from the model with CIs
table_model1_CI <-
  tbl_regression(ML_fit1, exponentiate = FALSE, intercept = TRUE) %>% 
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE) %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_header(statistic = "**T-Statistic**") # rename header from statistic to t-statistic
table_model1_CI
```

The adjusted R-squared is `r round(r2_1$model1, 3)`, and the RMSE is `r round(rmse_1$model1)`.

## Model 1 Performance

First, I will evaluate the model on the training data.

```{r, message=FALSE,  warning=FALSE}
# Training model 1
# Create a list to store R-squared and RMSE values
r2_1_train <- list()
rmse_1_train <- list()

# Model using recipes and workflows
# Define recipe for model
ML_recipe1_train <- recipe(Weightgrams ~ ModifiedColor, data = train_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable

# Specify model
ML_spec1_train <- linear_reg() %>% 
  set_engine("lm")

# Create work flow
ML_wf1_train <- workflow() %>%
  add_recipe(ML_recipe1_train) %>%
  add_model(ML_spec1_train)

# Fit the work flow
ML_fit1_train <- fit(ML_wf1_train, data = train_data)

# Summarize model fit with tidiers
ML_summary1_train <- summary(ML_fit1_train) # Summary of fit
tidy(ML_fit1_train, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit1_train) # Model-level statistics

# Produce predictions for R-squared
ML_fit1_pred_train <- predict(ML_fit1_train, new_data = train_data %>% select(-Weightgrams))
ML_fit1_pred_train

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit1_pred_train <- bind_cols(ML_fit1_pred_train, train_data %>% select(Weightgrams))
ML_fit1_pred_train

# Plot observed versus predicted
plot1_train <- ggplot(ML_fit1_pred_train, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot1_train

# Create metric set including RMSE and R-squared
metrics_fit1_train <- metric_set(rmse, rsq)
metrics_fit1_train(ML_fit1_pred_train, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit1_train <- as.data.frame(metrics_fit1_train(ML_fit1_pred_train, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit1_train

# Create a gtsummary table directly from the model
table_model1_train <-
  tbl_regression(ML_fit1_train, exponentiate = FALSE, intercept = TRUE, show_single_row = everything()) %>% 
  bold_p() %>% 
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE)
table_model1_train

# Extract R-squared and RMSE values for in-line text (because data frame with columns not list)
r2_1_train$model1_train <- metrics_fit1_train[metrics_fit1_train$.metric == "rsq", ".estimate"]
rmse_1_train$model1_train <- metrics_fit1_train[metrics_fit1_train$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model1_train, file = here::here("results", "tables", "table_model1_train.rds"))
```

The results are similar to those from the model run on all the data. The adjusted R-squared is `r round(r2_1_train$model1_train, 3)`, the same as the full model. The RMSE is `r round(rmse_1_train$model1_train)`, which is slightly lower than the full model, `r round(rmse_1$model1)`, which shows better performance.

Now, I will perform cross-validation on the training data.

```{r, message=FALSE,  warning=FALSE}
# CV model 1
# Set the seed for reproducibility
set.seed(rngseed)

# Create 10-fold cross-validation
folds <- vfold_cv(train_data, v = 10)
folds

# Create a list to store R-squared and RMSE values
r2_1_cv <- list()
rmse_1_cv <- list()

# Define recipe for model
ML_recipe1_cv <- recipe(Weightgrams ~ ModifiedColor, data = train_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable

# Set model specification
ML_spec1_cv <- linear_reg() %>% set_engine("lm")

# Create and initialize workflow that bundles together model specification and formula.
ML_wf1_cv <- workflow() %>% 
  add_model(ML_spec1_cv) %>% 
  add_recipe(ML_recipe1_cv)
  fit_resamples(ML_wf1_cv, resamples = folds)

# Resample the model
ML_wf1_cv_rs <- fit_resamples(ML_wf1_cv, resamples = folds)
ML_wf1_cv_rs

# Get model metrics
ML_cv1_metrics <- collect_metrics(ML_wf1_cv_rs)

# Filter the data frame to include only rows with "rmse" and "rsq" in the ".metric" column
ML_cv1_metrics <- ML_cv1_metrics[ML_cv1_metrics$.metric %in% c("rmse", "rsq"), !(names(ML_cv1_metrics) %in% ".config")]
ML_cv1_metrics

# Extract R-squared and RMSE values for in-line text
# For R-squared
r2_1_cv$model1_cv <- ML_cv1_metrics %>%
  filter(.metric == "rsq") %>%
  pull(mean)

# For RMSE
rmse_1_cv$model1_cv <- ML_cv1_metrics %>%
  filter(.metric == "rmse") %>%
  pull(mean)

# Save regression output into table
saveRDS(ML_cv1_metrics, file = here::here("results", "tables", "ML_cv1_metrics.rds"))
```
Performing cross-validation results in an RMSE that is the same as the training model `r round(rmse_1_cv$model1_cv)`, lower than the full model. The r-squared `r round(r2_1_cv$model1_cv, 3)` is also higher than the training `r round(r2_1_train$model1_train, 3)` and full models `r round(r2_1$model1, 3)`, demonstrating better model performance.

Finally, I'll run the model on the test data.

```{r, message=FALSE,  warning=FALSE}
# Testing model 1
# Create a list to store R-squared and RMSE values
r2_1_test <- list()
rmse_1_test <- list()

# Model using recipes and workflows
# Define recipe for model
ML_recipe1_test <- recipe(Weightgrams ~ ModifiedColor, data = test_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable

# Specify model
ML_spec1_test <- linear_reg() %>% 
  set_engine("lm")

# Create work flow
ML_wf1_test <- workflow() %>%
  add_recipe(ML_recipe1_test) %>%
  add_model(ML_spec1_test)

# Fit the work flow
ML_fit1_test <- fit(ML_wf1_test, data = test_data)

# Summarize model fit with tidiers
ML_summary1_test <- summary(ML_fit1_test) # Summary of fit
tidy(ML_fit1_test, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit1_test) # Model-level statistics

# Produce predictions for R-squared
ML_fit1_pred_test <- predict(ML_fit1_test, new_data = test_data %>% select(-Weightgrams))
ML_fit1_pred_test

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit1_pred_test <- bind_cols(ML_fit1_pred_test, test_data %>% select(Weightgrams))
ML_fit1_pred_test

# Plot observed versus predicted
plot1_test <- ggplot(ML_fit1_pred_test, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot1_test

# Create metric set including RMSE and R-squared
metrics_fit1_test <- metric_set(rmse, rsq)
metrics_fit1_test(ML_fit1_pred_test, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit1_test <- as.data.frame(metrics_fit1_test(ML_fit1_pred_test, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit1_test

# Create a gtsummary table directly from the model
table_model1_test <-
  tbl_regression(ML_fit1_test, exponentiate = FALSE, intercept = TRUE, show_single_row = everything()) %>% 
  bold_p() %>% 
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE)
table_model1_test

# Extract R-squared and RMSE values for in-line text (because data frame with columns not list)
r2_1_test$model1_test<- metrics_fit1_test[metrics_fit1_test$.metric == "rsq", ".estimate"]
rmse_1_test$model1_test <- metrics_fit1_test[metrics_fit1_test$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model1_test, file = here::here("results", "tables", "table_model1_test.rds"))
```

When running the model on the testing dataset, the results remain significant at the 0.05 level. The RMSE is `r round(rmse_1_test$model1_test)` and the r-squared is `r round(r2_1_test$model1_test, 3)`. This shows a slightly worse fit than the training results.

# Model 2

The second multiple linear regression model is as follows:

$$\text{BW} = \alpha + \beta_1 \text{MA} + \beta_2 \text{Gest} + \beta_3 \text{Nat} + \beta_4 \text{Age} + \epsilon$$

whereas `BW` is birth weight in grams, $\alpha$ is the average birth weight, $\beta_x$ are the group deviations from the average birth weight, `MA` is maternal ancestry (Euro-descent, Afro-descent), `Gest` is gestational status (nulliparous or multiparous), `Age` is maternal age in years, and $\epsilon$ is the error term. Results are displayed in Table 3.

```{r, message=FALSE,  warning=FALSE}
# Model 2: Final model for paper using all data
# Create a list to store R-squared values
r2_2 <- list()
rmse_2 <- list()

# Using recipes and workflows
# Define recipe for model
ML_recipe2 <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") #%>% # Set reference level for ModifiedColor
  #step_dummy(ModifiedColor) #%>% # Convert all nominal variables to dummy variables. Cannot do this if I am merging tables in gtsummary
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors. Cannot do this if I am merging tables in gtsummary

# Define model specification
ML_spec2 <- linear_reg() %>% 
  set_engine("lm") %>%
  set_mode("regression")

# Create workflow
ML_wf2 <- workflow() %>%
  add_recipe(ML_recipe2) %>%
  add_model(ML_spec2)

# Fit the workflow
ML_fit2 <- fit(ML_wf2, data = ML_linear)

# Summarize model fit with tidiers
ML_lm2_summary <- summary(ML_fit2) # Summary of fit
tidy(ML_fit2, conf.int = TRUE) %>% mutate(p.value = round(p.value, 5)) # Coefficients and confidence intervals with p-values rounded to 5 decimal places
glance(ML_fit2) # Model-level statistics

# Produce predictions for R-squared
ML_fit2_pred <- predict(ML_fit2, new_data = ML_linear %>% select(-Weightgrams))
ML_fit2_pred

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit2_pred <- bind_cols(ML_fit2_pred, ML_linear %>% select(Weightgrams))
ML_fit2_pred

# Plot observed versus predicted
plot2 <- ggplot(ML_fit2_pred, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot2

# Create metric set including RMSE and R-squared
metrics_fit2 <- metric_set(rmse, rsq)
metrics_fit2(ML_fit2_pred, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit2 <- as.data.frame(metrics_fit2(ML_fit2_pred, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit2

# Create a gtsummary table directly from the model
table_model2 <- tbl_regression(ML_fit2, exponentiate = FALSE, intercept = TRUE) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = TRUE, hide_p = TRUE) %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_header(statistic = "**T-Statistic**") # rename header from statistic to t-statistic
table_model2

# Extract RMSE and R2 for in-line text
r2_2$model2 <- metrics_fit2[metrics_fit2$.metric == "rsq", ".estimate"]
rmse_2$model2 <- metrics_fit2[metrics_fit2$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model2, file = here::here("results", "tables", "table_model2_final.rds"))

# Create a gtsummary table directly from the model with CIs
table_model2_CI <- tbl_regression(ML_fit2, exponentiate = FALSE, intercept = TRUE) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE) %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_header(statistic = "**T-Statistic**") # rename header from statistic to t-statistic
table_model2_CI
```

The RMSE is `r round(rmse_2$model2)` and the r-squared is `r round(r2_2$model2, 3)`.

## Model 2 Performance

Now, I'll run the model on the training data.

```{r, message=FALSE,  warning=FALSE}
# Training model 2
# Create a list to store R-squared and RMSE values
r2_2_train <- list()
rmse_2_train <- list()

# Using recipes and workflows
# Define recipe for model
ML_recipe2_train <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = train_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Define model specification
ML_spec2_train <- linear_reg() %>% 
  set_engine("lm") %>%
  set_mode("regression")

# Create workflow
ML_wf2_train <- workflow() %>%
  add_recipe(ML_recipe2_train) %>%
  add_model(ML_spec2_train)

# Fit the workflow
ML_fit2_train <- fit(ML_wf2_train, data = train_data)

# Summarize model fit with tidiers
ML_lm2_summary_train <- summary(ML_fit2_train) # Summary of fit
tidy(ML_fit2_train, conf.int = TRUE) %>% mutate(p.value = round(p.value, 5)) # Coefficients and confidence intervals with p-values rounded to 5 decimal places
glance(ML_fit2_train) # Model-level statistics

# Produce predictions for R-squared
ML_fit2_pred_train <- predict(ML_fit2_train, new_data = train_data %>% select(-Weightgrams))
ML_fit2_pred_train

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit2_pred_train <- bind_cols(ML_fit2_pred_train, train_data %>% select(Weightgrams))
ML_fit2_pred_train

# Plot observed versus predicted
plot2_train <- ggplot(ML_fit2_pred_train, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot2_train

# Create metric set including RMSE and R-squared
metrics_fit2_train <- metric_set(rmse, rsq)
metrics_fit2_train(ML_fit2_pred_train, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit2_train <- as.data.frame(metrics_fit2_train(ML_fit2_pred_train, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit2_train

# Create a gtsummary table directly from the model
table_model2_train <- pull_workflow_fit(ML_fit2_train) %>%
  tbl_regression(exponentiate = FALSE, intercept = TRUE) %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = TRUE, hide_p = TRUE) %>%
  bold_p()
table_model2_train

# Extract RMSE and R2 for in-line text
r2_2_train$model2_train <- metrics_fit2_train[metrics_fit2_train$.metric == "rsq", ".estimate"]
rmse_2_train$model2_train <- metrics_fit2_train[metrics_fit2_train$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model2_train, file = here::here("results", "tables", "table_model2_train.rds"))
```
The RMSE is `r round(rmse_2_train$model2_train)` and the r-squared is `r round(r2_2_train$model2_train, 3)` a slightly better performance than the full model.

Now I'll run cross-validation.

```{r, message=FALSE,  warning=FALSE}
# CV model 2
# Set the seed for reproducibility
set.seed(rngseed)

# Create 10-fold cross-validation
folds <- vfold_cv(train_data, v = 10)
folds

# Create a list to store R-squared and RMSE values
r2_2_cv <- list()
rmse_2_cv <- list()

# Define recipe for model
ML_recipe2_cv <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = train_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Set model specification
ML_spec2_cv <- linear_reg() %>% set_engine("lm")

# Create and initialize workflow that bundles together model specification and formula.
ML_wf2_cv <- workflow() %>% 
  add_model(ML_spec2_cv) %>% 
  add_recipe(ML_recipe2_cv)
  fit_resamples(ML_wf2_cv, resamples = folds)

# Resample the model
ML_wf2_cv_rs <- fit_resamples(ML_wf2_cv, resamples = folds)
ML_wf2_cv_rs

# Get model metrics
ML_cv2_metrics <- collect_metrics(ML_wf2_cv_rs)

# Filter the data frame to include only rows with "rmse" and "rsq" in the ".metric" column
ML_cv2_metrics <- ML_cv2_metrics[ML_cv2_metrics$.metric %in% c("rmse", "rsq"), !(names(ML_cv2_metrics) %in% ".config")]
ML_cv2_metrics

# Extract R-squared and RMSE values for in-line text
# For R-squared
r2_2_cv$model2_cv <- ML_cv2_metrics %>%
  filter(.metric == "rsq") %>%
  pull(mean)

# For RMSE
rmse_2_cv$model2_cv <- ML_cv2_metrics %>%
  filter(.metric == "rmse") %>%
  pull(mean)

# Save regression output into table
saveRDS(ML_cv2_metrics, file = here::here("results", "tables", "ML_cv2_metrics.rds"))
```
Here, the RMSE is `r round(rmse_2_cv$model2_cv)` and the r-squared is `r round(r2_2_cv$model2_cv, 3)`, the same as the training model `r round(rmse_2_train$model2_train)`, `r round(r2_2_train$model2_train, 3)`.

Finally, I'll run the model on the test data.

```{r, message=FALSE,  warning=FALSE}
# Testing model 2
# Create a list to store R-squared and RMSE values
r2_2_test <- list()
rmse_2_test <- list()

# Using recipes and workflows
# Define recipe for model
ML_recipe2_test <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = test_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Define model specification
ML_spec2_test <- linear_reg() %>% 
  set_engine("lm") %>%
  set_mode("regression")

# Create workflow
ML_wf2_test <- workflow() %>%
  add_recipe(ML_recipe2_test) %>%
  add_model(ML_spec2_test)

# Fit the workflow
ML_fit2_test <- fit(ML_wf2_test, data = test_data)

# Summarize model fit with tidiers
ML_lm2_summary_test <- summary(ML_fit2_test) # Summary of fit
tidy(ML_fit2_test, conf.int = TRUE) %>% mutate(p.value = round(p.value, 5)) # Coefficients and confidence intervals with p-values rounded to 5 decimal places
glance(ML_fit2_test) # Model-level statistics

# Produce predictions for R-squared
ML_fit2_pred_test <- predict(ML_fit2_test, new_data = test_data %>% select(-Weightgrams))
ML_fit2_pred_test

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit2_pred_test <- bind_cols(ML_fit2_pred_test, test_data %>% select(Weightgrams))
ML_fit2_pred_test

# Plot observed versus predicted
plot2_test <- ggplot(ML_fit2_pred_test, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot2_test

# Create metric set including RMSE and R-squared
metrics_fit2_test <- metric_set(rmse, rsq)
metrics_fit2_test(ML_fit2_pred_test, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit2_test <- as.data.frame(metrics_fit2_test(ML_fit2_pred_test, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit2_test

# Create a gtsummary table directly from the model
table_model2_test <- pull_workflow_fit(ML_fit2_test) %>%
  tbl_regression(exponentiate = FALSE, intercept = TRUE) %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = TRUE, hide_p = TRUE) %>%
  bold_p()
table_model2_test

# Extract R-squared and RMSE values for in-line text (because data frame with columns not list)
r2_2_test$model2_test<- metrics_fit2_test[metrics_fit2_test$.metric == "rsq", ".estimate"]
rmse_2_test$model2_test <- metrics_fit2_test[metrics_fit2_test$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model2_test, file = here::here("results", "tables", "table_model2_test.rds"))
```

Running the model on the test data, there are no significant results at any alpha level.

## Models 1 and 2 Sub-Samples

```{r, message=FALSE,  warning=FALSE}
# Model 1: Sub-sample simple model (ML_lm1_sub)
# Create a list to store R-squared values
r_squared_values_sub <- list()

# Filter data based on desired conditions
filtered_data <- ML_linear %>%
  filter(Brazilian == "Brazilian")

# Set reference level for "Euro-descent"
ML_linear$ModifiedColor <- relevel(ML_linear$ModifiedColor, ref = "Euro-Descent")

# Specify and fit model using linear_reg() (default engine, OLS) from tidymodels
ML_lm1_sub <- linear_reg() %>% set_engine("lm") %>% fit(Weightgrams ~ ModifiedColor, data = filtered_data)

# Summarize model fit with tidiers
ML_lm1_sub_summary <- summary(ML_lm1_sub) # Summary of fit
tidy(ML_lm1_sub, conf.int = TRUE) %>% mutate(p.value = round(p.value, 5)) # Coefficients and confidence intervals with p-values rounded to 5 decimal places
ML_lm1_sub_stat <- glance(ML_lm1_sub) # Extract model-level statistics

# Extract R-squared value
r_squared_values_sub$ML_lm1_sub <- ML_lm1_sub_stat$r.squared

# Create a gtsummary table directly from the model
ML_lm1_sub_table <- 
  tbl_regression(ML_lm1_sub, exponentiate = FALSE, intercept = TRUE) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = TRUE, hide_p = TRUE) %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_header(statistic = "**T-Statistic**") # rename header from statistic to t-statistic
ML_lm1_sub_table

# Save regression output into table
saveRDS(ML_lm1_sub_table, file = here::here("results", "tables", "ML_lm1_sub_table.rds"))
```
```{r, message=FALSE,  warning=FALSE}
# Model 2: Sub-sample multivariate model (ML_lm2_sub)
# Filter data based on desired conditions
filtered_data <- ML_linear %>%
  filter(Brazilian == "Brazilian")

# Set reference level for "Euro-descent"
ML_linear$ModifiedColor <- relevel(ML_linear$ModifiedColor, ref = "Euro-Descent")

# Specify and fit model using linear_reg() (default engine, OLS) from tidymodels
ML_lm2_sub <- linear_reg() %>% set_engine("lm") %>% fit(Weightgrams ~ ModifiedColor + Age + + ModifiedStatus, data = filtered_data)

# Summarize model fit with tidiers
ML_lm2_sub_summary <- summary(ML_lm2_sub) # Summary of fit
tidy(ML_lm2_sub, conf.int = TRUE) %>% mutate(p.value = round(p.value, 5)) # Coefficients and confidence intervals with p-values rounded to 5 decimal places
ML_lm2_sub_stat <- glance(ML_lm2_sub) # Extract model-level statistics

# Extract R-squared value
r_squared_values_sub$ML_lm2_sub <- ML_lm2_sub_stat$r.squared

# Create a gtsummary table directly from the model
ML_lm2_sub_table <- 
  tbl_regression(ML_lm2_sub, exponentiate = FALSE, intercept = TRUE) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = TRUE, hide_p = TRUE) %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_header(statistic = "**T-Statistic**") # rename header from statistic to t-statistic
ML_lm2_sub_table

# Save regression output into table
saveRDS(ML_lm2_sub_table, file = here::here("results", "tables", "ML_lm2_sub_table.rds"))
```

# Merging Model 1 and 2 and sub-sample 1 and 2 into Table

```{r, message=FALSE,  warning=FALSE}
# Combine the four tables using tbl_merge
table3 <- tbl_merge(
  list(table_model1, ML_lm1_sub_table, table_model2, ML_lm2_sub_table),
  tab_spanner = c("**Model 1**", "**Sub-Model 1**", "**Model 2**", "**Sub-Model 2**")) %>%
  modify_table_body(fun = ~.x %>% arrange(variable))
table3

# Save regression output into table
saveRDS(table3, file = here::here("results", "tables", "table3_final.rds"))
```

# Model 3

The second two models are also simple and multiple linear regressions with outcome variable as birthweight in grams. However, in these two models, I maintained maternal skin color as three categories, White (the reference group), Black, and mixed race to see if there is differential outcomes for Black and mixed-race women. The first simple linear regression model is as follows:

$$\text{BW} = \alpha + \beta_1 \text{MC} + \epsilon$$
whereas $\alpha$ is the average birth weight, $\beta_1$ is the group deviation from the average birth weight, `BW` is birth weight in grams, `MC` is maternal skin color (White, Black, and mixed race), and $\epsilon$ is the error term. Results are displayed in Table 4.

```{r, message=FALSE,  warning=FALSE}
# Model 3: Final model for paper using all data
# Create a list to store R-squared and RMSE values
r2_3 <- list()
rmse_3 <- list()

# Using recipes and workflows
# Simple linear model
# Specify recipe
ML_recipe3 <- recipe(Weightgrams ~ Color, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White") #%>% # Set reference level for White
  #step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Specify model
ML_spec3 <- linear_reg() %>%
  set_engine("lm")

# Create workflow
ML_wf3 <- workflow() %>%
  add_recipe(ML_recipe3) %>%
  add_model(ML_spec3)

# Fit the workflow
ML_fit3 <- fit(ML_wf3, data = ML_linear)

# Summarize model fit with tidiers and create a gtsummary table directly from the model
ML_summary3 <- summary(ML_fit3)
tidy(ML_fit3, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit3) # Model-level statistics

# Produce predictions for R-squared
ML_fit3_pred <- predict(ML_fit3, new_data = ML_linear %>% select(-Weightgrams))
ML_fit3_pred

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit3_pred <- bind_cols(ML_fit3_pred, ML_linear %>% select(Weightgrams))
ML_fit3_pred

# Plot observed versus predicted
plot3 <- ggplot(ML_fit3_pred, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot3

# Create metric set including RMSE and R-squared
metrics_fit3 <- metric_set(rmse, rsq)
metrics_fit3(ML_fit3_pred, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit3 <- as.data.frame(metrics_fit3(ML_fit3_pred, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit3

# Create a gtsummary table from model
table_model3 <- tbl_regression(ML_fit3, exponentiate = FALSE, intercept = TRUE) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = TRUE, hide_p = TRUE) %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_header(statistic = "**T-Statistic**") # rename header from statistic to t-statistic
table_model3

# Extract RMSE and R2 for in-line text
r2_3$model3 <- metrics_fit3[metrics_fit3$.metric == "rsq", ".estimate"]
rmse_3$model3 <- metrics_fit3[metrics_fit3$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model3, file = here("results", "tables", "table_model3_final.rds"))

# Create a gtsummary table from model with CIs
table_model3_CI <- tbl_regression(ML_fit3, exponentiate = FALSE, intercept = TRUE) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE) %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_header(statistic = "**T-Statistic**") # rename header from statistic to t-statistic
table_model3_CI
```

The RMSE is `r round(rmse_3$model3)` and the r-squared is `r round(r2_3$model3, 3)`.

## Model 3 Performance

Now I'll evaluate the model on our training data.

```{r, message=FALSE,  warning=FALSE}
# Training model 3
# Create a list to store R-squared and RMSE values
r2_3_train <- list()
rmse_3_train <- list()

# Using recipes and workflows
# Specify recipe
ML_recipe3_train <- recipe(Weightgrams ~ Color, data = train_data) %>%
  step_relevel(Color, ref_level = "White")

# Specify model
ML_spec3_train <- linear_reg() %>%
  set_engine("lm")

# Create workflow
ML_wf3_train <- workflow() %>%
  add_recipe(ML_recipe3_train) %>%
  add_model(ML_spec3_train)

# Fit the workflow
ML_fit3_train <- fit(ML_wf3_train, data = train_data)

# Summarize model fit with tidiers and create a gtsummary table directly from the model
ML_summary3_train <- summary(ML_fit3_train)
tidy(ML_fit3_train, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit3_train) # Model-level statistics

# Produce predictions for R-squared
ML_fit3_pred_train <- predict(ML_fit3_train, new_data = train_data %>% select(-Weightgrams))
ML_fit3_pred_train

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit3_pred_train <- bind_cols(ML_fit3_pred_train, train_data %>% select(Weightgrams))
ML_fit3_pred_train

# Plot observed versus predicted
plot3_train <- ggplot(ML_fit3_pred_train, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot3_train

# Create metric set including RMSE and R-squared
metrics_fit3_train <- metric_set(rmse, rsq)
metrics_fit3_train(ML_fit3_pred_train, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit3_train <- as.data.frame(metrics_fit3_train(ML_fit3_pred_train, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit3_train

# Create a gtsummary table from model
table_model3_train <- tbl_regression(ML_fit3_train, exponentiate = FALSE, intercept = TRUE) %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = TRUE, hide_p = TRUE)
table_model3_train

# Extract RMSE and R2 for in-line text
r2_3_train$model3_train <- metrics_fit3_train[metrics_fit3_train$.metric == "rsq", ".estimate"]
rmse_3_train$model3_train <- metrics_fit3_train[metrics_fit3_train$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model3_train, file = here("results", "tables", "table_model3_train.rds"))
```

The RMSE is `r round(rmse_3_train$model3_train)` and the r-squared is `r round(r2_3_train$model3_train, 3)`, slightly better performance than the full model `r round(r2_3$model3, 3)`. The results are not significantly different from the full model.

Now I'll run cross-validation.

```{r, message=FALSE,  warning=FALSE}
# CV model 3
# Set the seed for reproducibility
set.seed(rngseed)

# Create 10-fold cross-validation
folds <- vfold_cv(train_data, v = 10)
folds

# Create list to store R-squared and RMSE values
r2_3_cv <- list()
rmse_3_cv <- list()

# Define recipe for model
ML_recipe3_cv <- recipe(Weightgrams ~ Color, data = train_data) %>%
  step_relevel(Color, ref_level = "White")

# Set model specification
ML_spec3_cv <- linear_reg() %>% set_engine("lm")

# Create and initialize workflow that bundles together model specification and formula.
ML_wf3_cv <- workflow() %>% 
  add_model(ML_spec3_cv) %>% 
  add_recipe(ML_recipe3_cv)
  fit_resamples(ML_wf3_cv, resamples = folds)

# Resample the model
ML_wf3_cv_rs <- fit_resamples(ML_wf3_cv, resamples = folds)
ML_wf3_cv_rs

# Get model metrics
collect_metrics(ML_wf3_cv_rs)

# Get model metrics
ML_cv3_metrics <- collect_metrics(ML_wf3_cv_rs)

# Filter the data frame to include only rows with "rmse" and "rsq" in the ".metric" column
ML_cv3_metrics <- ML_cv3_metrics[ML_cv3_metrics$.metric %in% c("rmse", "rsq"), !(names(ML_cv3_metrics) %in% ".config")]
ML_cv3_metrics

# Extract R-squared and RMSE values for in-line text
# For R-squared
r2_3_cv$model3_cv <- ML_cv3_metrics %>%
  filter(.metric == "rsq") %>%
  pull(mean)

# For RMSE
rmse_3_cv$model3_cv <- ML_cv3_metrics %>%
  filter(.metric == "rmse") %>%
  pull(mean)

# Save regression output into table
saveRDS(ML_cv3_metrics, file = here::here("results", "tables", "ML_cv3_metrics.rds"))
```
Using cross-validation, the RMSE is `r round(rmse_3_cv$model3_cv)` and the r-squared is `r round(r2_3_cv$model3_cv, 3)`, similar performances to the training model `r round(rmse_3_train$model3_train)`, `r round(r2_3_train$model3_train, 3)`.

Finally, I'll run the model on the test data.

```{r, message=FALSE, warning=FALSE}
# Testing model 3
# Create a list to store R-squared and RMSE values
r2_3_test <- list()
rmse_3_test <- list()

# Using recipes and workflows
# Specify recipe
ML_recipe3_test <- recipe(Weightgrams ~ Color, data = test_data) %>%
  step_relevel(Color, ref_level = "White")

# Specify model
ML_spec3_test <- linear_reg() %>%
  set_engine("lm")

# Create workflow
ML_wf3_test <- workflow() %>%
  add_recipe(ML_recipe3_test) %>%
  add_model(ML_spec3_test)

# Fit the workflow
ML_fit3_test <- fit(ML_wf3_test, data = test_data)

# Summarize model fit with tidiers and create a gtsummary table directly from the model
ML_summary3_test <- summary(ML_fit3_test)
tidy(ML_fit3_test, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit3_test) # Model-level statistics

# Produce predictions for R-squared
ML_fit3_pred_test <- predict(ML_fit3_test, new_data = test_data %>% select(-Weightgrams))
ML_fit3_pred_test

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit3_pred_test <- bind_cols(ML_fit3_pred_test, test_data %>% select(Weightgrams))
ML_fit3_pred_test

# Plot observed versus predicted
plot3_test <- ggplot(ML_fit3_pred_test, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot3_test

# Create metric set including RMSE and R-squared
metrics_fit3_test <- metric_set(rmse, rsq)
metrics_fit3_test(ML_fit3_pred_test, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit3_test <- as.data.frame(metrics_fit3_test(ML_fit3_pred_test, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit3_test

# Create a gtsummary table from model
table_model3_test <- tbl_regression(ML_fit3_test, exponentiate = FALSE, intercept = TRUE) %>%
  bold_p() %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE)
table_model3_test

# Extract R-squared and RMSE values for in-line text (because data frame with columns not list)
r2_3_test$model3_test<- metrics_fit3_test[metrics_fit3_test$.metric == "rsq", ".estimate"]
rmse_3_test$model3_test <- metrics_fit3_test[metrics_fit3_test$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model3_test, file = here("results", "tables", "table_model3_test.rds"))
```

Much like in the other testing models, the results are not significant (except the intercept). The performance is also worse (RMSE of `r round(rmse_3_test$model3_test)` and r-squared of `r round(r2_3_test$model3_test, 3)`).

# Model 4

The second multiple linear regression model is as follows:

$$\text{BW} = \alpha + \beta_1 \text{MC} + \beta_2 \text{Gest} + \beta_3 \text{Nat} + \beta_4 \text{Age} + \epsilon$$

whereas $\alpha$ is the average birth weight, $\beta_x$ are the group deviations from the average birth weight, `BW` is birth weight in grams, `MC` is maternal skin color (White, Black, and mixed race), `Gest` is gestational status (nulliparous or multiparous), `Nat` is maternal nationality (Brazilian, Latin American, European, Middle Eastern), `Age` is maternal age in years, and $\epsilon$ is the error term.

```{r, message=FALSE,  warning=FALSE}
# Model 4: Final model for paper using all data
# Create a list to store R-squared values
r2_4 <- list()
rmse_4 <- list()

# Using recipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4 <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") #%>% # Set reference level for ModifiedStatus
  #step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Specify model
ML_spec4 <- linear_reg() %>%
  set_engine("lm")

# Create workflow
ML_wf4 <- workflow() %>%
  add_recipe(ML_recipe4) %>%
  add_model(ML_spec4)

# Fit the workflow
ML_fit4 <- fit(ML_wf4, data = ML_linear)

# Summarize model fit with tidiers and create a gtsummary table directly from the model
ML_summary4 <- summary(ML_fit4)
tidy(ML_fit4, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit4) # Model-level statistics

# Produce predictions for R-squared
ML_fit4_pred <- predict(ML_fit4, new_data = ML_linear %>% select(-Weightgrams))
ML_fit4_pred

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit4_pred <- bind_cols(ML_fit4_pred, ML_linear %>% select(Weightgrams))
ML_fit4_pred

# Plot observed versus predicted
plot4 <- ggplot(ML_fit4_pred, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot4

# Create metric set including RMSE and R-squared
metrics_fit4 <- metric_set(rmse, rsq)
metrics_fit4(ML_fit4_pred, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit4 <- as.data.frame(metrics_fit4(ML_fit4_pred, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit4

# Create a gtsummary table from model
table_model4 <- tbl_regression(ML_fit4, exponentiate = FALSE, intercept = TRUE) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = TRUE, hide_p = TRUE) %>% # Hide confidence intervals and p-values
  add_glance_table(include = c(r.squared, nobs)) %>% # Add R-squared and number of observations
  modify_header(statistic = "**T-Statistic**")
table_model4

# Extract RMSE and R2 for in-line text
r2_4$model4 <- metrics_fit4[metrics_fit4$.metric == "rsq", ".estimate"]
rmse_4$model4 <- metrics_fit4[metrics_fit4$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model4, file = here("results", "tables", "table_model4_final.rds"))

# Create a gtsummary table from model with CIs
table_model4_CI <- tbl_regression(ML_fit4, exponentiate = FALSE, intercept = TRUE) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE) %>% # Hide confidence intervals and p-values
  add_glance_table(include = c(r.squared, nobs)) %>% # Add R-squared and number of observations
  modify_header(statistic = "**T-Statistic**")
table_model4_CI
```
The RMSE is `r round(rmse_4$model4)` and the r-squared is `r round(r2_4$model4, 3)`.

## Model 4 Performance

I will now evaluate the model on the training data.

```{r, message=FALSE,  warning=FALSE}
# Model 4 training
# Create list to store R-squared and RMSE values
r2_4_train <- list()
rmse_4_train <- list()

# Using recipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4_train <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = train_data) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") %>% # Set reference level for ModifiedStatus
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Specify model
ML_spec4_train <- linear_reg() %>%
  set_engine("lm")

# Create workflow
ML_wf4_train <- workflow() %>%
  add_recipe(ML_recipe4_train) %>%
  add_model(ML_spec4_train)

# Fit the workflow
ML_fit4_train <- fit(ML_wf4_train, data = train_data)

# Summarize model fit with tidiers and create a gtsummary table directly from the model
ML_summary4_train <- summary(ML_fit4_train)
tidy(ML_fit4_train, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit4_train) # Model-level statistics

# Produce predictions for R-squared
ML_fit4_pred_train <- predict(ML_fit4_train, new_data = train_data %>% select(-Weightgrams))
ML_fit4_pred_train

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit4_pred_train <- bind_cols(ML_fit4_pred_train, train_data %>% select(Weightgrams))
ML_fit4_pred_train

# Plot observed versus predicted
plot4_train <- ggplot(ML_fit4_pred_train, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot4_train

# Create metric set including RMSE and R-squared
metrics_fit4_train <- metric_set(rmse, rsq)
metrics_fit4_train(ML_fit4_pred_train, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit4_train <- as.data.frame(metrics_fit4_train(ML_fit4_pred_train, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit4_train

# Create a gtsummary table from model
table_model4_train <- tbl_regression(ML_fit4_train, exponentiate = FALSE, intercept = TRUE) %>%
  bold_p() %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE)
table_model4_train

# Extract RMSE and R2 for in-line text
r2_4_train$model4_train <- metrics_fit4_train[metrics_fit4_train$.metric == "rsq", ".estimate"]
rmse_4_train$model4_train <- metrics_fit4_train[metrics_fit4_train$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model4_train, file = here("results", "tables", "table_model4_train.rds"))
```

The RMSE is `r round(rmse_4_train$model4_train)` and the r-squared is `r round(r2_4_train$model4_train, 3)` a slightly better performance than the full model `r round(r2_4$model4, 3)`.

Now I'll run cross-validation.

```{r, message=FALSE,  warning=FALSE}
# CV model 4
# Set the seed for reproducibility
set.seed(rngseed)

# Create 10-fold cross-validation
folds <- vfold_cv(train_data, v = 10)
folds

# Create list to store R-squared and RMSE values
r2_4_cv <- list()
rmse_4_cv <- list()

# Define recipe for model
ML_recipe4_cv <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") %>% # Set reference level for ModifiedStatus
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Set model specification
ML_spec4_cv <- linear_reg() %>% set_engine("lm")

# Create and initialize workflow that bundles together model specification and formula.
ML_wf4_cv <- workflow() %>% 
  add_model(ML_spec4_cv) %>% 
  add_recipe(ML_recipe4_cv)
  fit_resamples(ML_wf4_cv, resamples = folds)

# Resample the model
ML_wf4_cv_rs <- fit_resamples(ML_wf4_cv, resamples = folds)
ML_wf4_cv_rs

# Get model metrics
collect_metrics(ML_wf4_cv_rs)

# Get model metrics
ML_cv4_metrics <- collect_metrics(ML_wf4_cv_rs)

# Filter the data frame to include only rows with "rmse" and "rsq" in the ".metric" column
ML_cv4_metrics <- ML_cv4_metrics[ML_cv4_metrics$.metric %in% c("rmse", "rsq"), !(names(ML_cv4_metrics) %in% ".config")]
ML_cv4_metrics

# Extract R-squared and RMSE values for in-line text
# For R-squared
r2_4_cv$model4_cv <- ML_cv4_metrics %>%
  filter(.metric == "rsq") %>%
  pull(mean)

# For RMSE
rmse_4_cv$model4_cv <- ML_cv4_metrics %>%
  filter(.metric == "rmse") %>%
  pull(mean)

# Save regression output into table
saveRDS(ML_cv4_metrics, file = here::here("results", "tables", "ML_cv4_metrics.rds"))
``` 
Here, the RMSE is `r round(rmse_4_cv$model4_cv)` and the r-squared is `r round(r2_4_cv$model4_cv, 3)`, a better performance than both the full model `r round(rmse_4$model4)`, `r round(r2_4$model4, 3)` and the training model `r round(r2_4_train$model4_train, 3)`, `r round(rmse_4_train$model4_train)`.

Finally, I will run model on the test data.

```{r, message=FALSE, warning=FALSE}
# Model 4 testing
# Create list to store R-squared and RMSE values
r2_4_test <- list()
rmse_4_test <- list()

# Using recipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4_test <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = test_data) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") %>% # Set reference level for ModifiedStatus
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

# Specify model
ML_spec4_test <- linear_reg() %>%
  set_engine("lm")

# Create workflow
ML_wf4_test <- workflow() %>%
  add_recipe(ML_recipe4_test) %>%
  add_model(ML_spec4_test)

# Fit the workflow
ML_fit4_test <- fit(ML_wf4_test, data = test_data)

# Summarize model fit with tidiers and create a gtsummary table directly from the model
ML_summary4_test <- summary(ML_fit4_test)
tidy(ML_fit4_test, conf.int = TRUE) # Coefficients and confidence intervals
glance(ML_fit4_test) # Model-level statistics

# Produce predictions for R-squared
ML_fit4_pred_test <- predict(ML_fit4_test, new_data = test_data %>% select(-Weightgrams))
ML_fit4_pred_test

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit4_pred_test <- bind_cols(ML_fit4_pred_test, test_data %>% select(Weightgrams))
ML_fit4_pred_test

# Plot observed versus predicted
plot4_test <- ggplot(ML_fit4_pred_test, aes(x = Weightgrams, y = .pred)) +
  geom_abline(lty = 2) + # Add a dashed line to represent the 1:1 line
  geom_point(alpha = 0.5) +
  labs(x = "Observed", y = "Predicted") + #Scale and size x- and y-axis uniformly
  coord_obs_pred()
plot4_test

# Create metric set including RMSE and R-squared
metrics_fit4_test <- metric_set(rmse, rsq)
metrics_fit4_test(ML_fit4_pred_test, truth = Weightgrams, estimate = .pred)

# Convert the output to a data frame and mutate the .estimate column to numeric
metrics_fit4_test <- as.data.frame(metrics_fit4_test(ML_fit4_pred_test, truth = Weightgrams, estimate = .pred)) %>%
  mutate(.estimate = as.numeric(.estimate))
options(scipen = 999) # set opens to avoid scientific notation
metrics_fit4_test

# Create a gtsummary table from model
table_model4_test <- tbl_regression(ML_fit4_test, exponentiate = FALSE, intercept = TRUE) %>%
  bold_p() %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE)
table_model4_test

# Extract R-squared and RMSE values for in-line text (because data frame with columns not list)
r2_4_test$model4_test<- metrics_fit4_test[metrics_fit4_test$.metric == "rsq", ".estimate"]
rmse_4_test$model4_test <- metrics_fit4_test[metrics_fit4_test$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model4_test, file = here("results", "tables", "table_model4_test.rds"))
```

Again, most results are not significant except the intercept given the small sample size.

## Merge Models 3 and 4

```{r, message=FALSE,  warning=FALSE}
# Combine the two tables using tbl_merge
table4 <- tbl_merge(list(table_model3, table_model4), tab_spanner = c("**Model 3**", "**Model 4**")) %>%
  modify_table_body(fun = ~.x %>% arrange(variable))
table4

# Save regression output into table
saveRDS(table4, file = here("results", "tables", "table4_final.rds"))
```